---
title: "20241030_2300_virosporequant_magalie"
author: "Joy O'Brien"
date: "2024-10-31"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)  
library(vegan)
library(ggplot2)
library(mgcv)

# Read data into 
viro_dist <- read_excel("20241030_2300_virosporequant_magalie.xlsx", sheet = "Virospore_quant_orig")
```

```{r}
# Since the sampling points are fixed values
viro_dist$Distance_from_PFU_cm <- factor(viro_dist$Distance_from_PFU_cm)

# Boxplot of recovered VPFU by distance from PFU
ggplot(viro_dist, aes(x = Distance_from_PFU_cm, y = Recovered_viral_PFU, fill = Distance_from_PFU_cm)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Recovered VPFU by Distance from PFU", x = "Distance from PFU", y = "Recovered VPFU") +
  theme_minimal()

# ANOVA
anova_result <- aov(Recovered_VPFU ~ Distance_from_PFU_cm, data = viro_dist)
summary(anova_result)

```
```{r}
# Playing with the data to see if it fits any known distribution
# Looking to fit the data, normal distribution
library(fitdistrplus)
fit_normal <-fitdist(viro_dist$Recovered_viral_PFU, "norm")
summary(fit_normal)
plot(fit_normal)

AIC_norm <- AIC(fit_normal)

# Fitting to exponential
fit_expon <-fitdist(viro_dist$Recovered_viral_PFU, "exp")
summary(fit_expon)
plot(fit_expon)
AIC_expon <- AIC(fit_expon)

# Fitting to poisson 
fit_poisson <-fitdist(viro_dist$Recovered_viral_PFU, "pois")
summary(fit_poisson)
plot(fit_poisson)
AIC_poisson <- AIC(fit_poisson)

# Fitting to gamma? I have 0s though...
library(MASS)
fit_gam <- fitdistr(viro_dist$Recovered_viral_PFU, "gamma")
summary(fit_gam)
plot(fit_gam)

# Fitting to beta
library(fitdistrplus)
beta_fit <- fitdist(viro_dist$Recovered_viral_PFU / max(viro_dist$Recovered_vpfu), "beta")

```
```{r}
# Continuing to play with the data
#Log transform the data and fit to normal distribution
viro_dist$log_recovered_VPFU <- log(viro_dist$Recovered_VPFU + 1)
library(fitdistrplus)
fit_normal_log <-fitdist(viro_dist$log_recovered_VPFU, "norm")
summary(fit_normal_log)
plot(fit_normal_log)

AIC_norm_log <- AIC(fit_normal_log)

# Log 10 transform the data and fit to normal distribution
viro_dist$log10_recovered_VPFU <- log10(viro_dist$Recovered_VPFU + 1)
library(fitdistrplus)
fit_normal_log10 <-fitdist(viro_dist$log10_recovered_VPFU, "norm")
summary(fit_normal_log10)
plot(fit_normal_log10)

AIC_norm_log10 <- AIC(fit_normal_log10)
```

```{r}
# Based on the lowest AIC score, I think that log10 transformation of the data is the best choice??

summary_datalog10 <- viro_dist %>%
  group_by(Distance_from_PFU_cm) %>%
  summarise(
    mean_VPFU = mean(log10_recovered_VPFU, na.rm = TRUE),
    std_dev = sd(log10_recovered_VPFU, na.rm = TRUE)
  ) %>%
  mutate(
    lower_bound = mean_VPFU - std_dev,
    upper_bound = mean_VPFU + std_dev
  )

# Plot the data
ggplot(summary_datalog10, aes(x = Distance_from_PFU_cm)) +
  geom_bar(aes(y = mean_VPFU), stat = "identity", fill = "lightgreen", color = "green", width = 0.7) +
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2, color = "green") +
  labs(x = "Region", y = "Recovered PFU") +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = -0.1)
  ) +
  ggtitle("Recovered PFU by Plaque Region")
```

```{r}
library(ggplot2)
library(dplyr)

# Normalize the data to the area before calculating summary statistics
viro_dist$Recovered_VPFU_per_cm2 <- viro_dist$Recovered_viral_PFU / 0.025
# Calculate mean and standard deviation of recovered PFU for each sample region
summary_data <- viro_dist %>%
  group_by(Distance_from_PFU_cm) %>%
  summarise(
    mean_VPFU = mean(Recovered_VPFU_per_cm2, na.rm = TRUE),
    std_dev = sd(Recovered_VPFU_per_cm2, na.rm = TRUE)
  ) %>%
  mutate(
    lower_bound = mean_VPFU - std_dev,
    upper_bound = mean_VPFU + std_dev
  )

# Plot the standardized data
ggplot(viro_dist, aes(x = Distance_from_PFU_cm, y = Recovered_VPFU_per_cm2)) +
  geom_point() +
  labs(y = "Recovered VPFU per cm²")


# Plot the data
ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  geom_bar(aes(y = mean_VPFU), stat = "identity", fill = "lightgreen", color = "green", width = 0.7) +
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2, color = "green") +
  labs(x = "Region", y = "Recovered PFU") +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = -0.1)
  ) +
  ggtitle("Recovered PFU by Plaque Region")

```
```{r}
# Plot including the raw data-- per mm2
# Plot including raw data 
ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  geom_point(data = viro_dist, aes(y = Recovered_VPFU_per_cm2), color = "black", alpha = 0.6, size = 1.5) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = FALSE) +
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), size = 1) +
  geom_smooth(aes(y = lower_bound, color = "Standard deviation"), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), linetype = "dashed") +
  # Customizing colors to match Magalie
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "green")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospore PFU/cm²") +
  theme_minimal(base_size = 14) +
  theme(legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1)
  )
```

```{r}
# Plot including raw data 
ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  geom_point(data = viro_dist, aes(y = Recovered_viral_PFU), color = "black", alpha = 0.6, size = 1.5) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = FALSE) +
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), size = 1) +
  geom_smooth(aes(y = lower_bound, color = "Standard deviation"), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), linetype = "dashed") +
  # Customizing colors to match Magalie
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "green")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  # Theme adjustments for a clean look
  theme_minimal(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1)
  )
```



```{r}
library(ggplot2)

# Plot using geom_smooth with LOESS smoothing
ggplot(summary_data, aes(x = Distance_from_PFU_cm, y = mean_VPFU)) +
  # Smoothed line with ribbon for standard deviation
  geom_smooth(aes(color = "Mean", fill = "Standard deviation"), 
              method = "loess", span = 0.3, se = TRUE, 
              size = 1, linetype = "solid") +
  # Customizing colors to match the aesthetic
  scale_color_manual(values = c("Mean" = "green")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from center (µm)", y = "Recovered PFU") +
  theme_minimal(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1)
  ) +
  ggtitle("C")

```

```{r}

library(ggplot2)

# Plot using geom_smooth with LOESS smoothing
ggplot(summary_data, aes(x = Distance_from_PFU_cm, y = mean_VPFU)) +
  # Smoothed line with ribbon for standard deviation
  geom_smooth(aes(color = "Mean", fill = "Standard deviation"), 
              method = "loess", span = 0.3, se = TRUE, 
              size = 1, linetype = "solid") +
  # Customizing colors to match the aesthetic
  scale_color_manual(values = c("Mean" = "green")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  # Theme adjustments for a clean look
  theme_minimal(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1)
  )

```
```{r}
# Informed by the MATLAB code from Magalie et al
library(ggplot2)

# Assuming summary_data is your data frame and it contains the necessary columns
ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  # Shaded area representing ± 1 standard deviation
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = FALSE) +
  # Solid line for the mean intensity
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), size = 1) +
  # Dashed lines for ± 1 standard deviation
  geom_smooth(aes(y = lower_bound, color = "Standard deviation"), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), linetype = "dashed") +
  # Customizing colors to match the aesthetic
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "green")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  # Theme adjustments for a clean look
  theme_minimal(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1)
  )

# Getting lots of errors with this code probably because the span parameter is too small, or when there are fewer datapoints than DoF ; so we need to increase the span and use a different method for smoothing

```



```{r setup, include=FALSE}
library(ggplot2)
library

ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  # Shaded area representing ± 1 standard deviation
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = FALSE) +
  # Solid line for the mean intensity
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), method = "gam", formula = y ~ s(x, bs = "cs"), size = 1) +
  # Dashed lines for ± 1 standard deviation
  geom_smooth(aes(y = lower_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs"), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs"), linetype = "dashed") +
  # Customizing colors to match the aesthetic
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "green")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  # Theme adjustments for a clean look
  theme_minimal(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1)
  )
```


```{r setup, include=FALSE}
library(ggplot2)
library(mgcv) # For GAM smoothing

# Assuming 'summary_data' contains columns: 'Distance_from_PFU_cm', 'mean_VPFU', 'lower_bound', 'upper_bound'
ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  # Shaded area representing ± 1 standard deviation
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = FALSE) +
  # Solid line for the mean intensity
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), method = "gam", formula = y ~ s(x, bs = "cs"), linewidth = 1) +
  # Dashed lines for ± 1 standard deviation
  geom_smooth(aes(y = lower_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs"), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs"), linetype = "dashed") +
  # Customizing colors to match the aesthetic
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "darkgreen")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  # Theme adjustments for a clean look
  theme_minimal(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 14, face = "bold", hjust = -0.1))
```


```{r setup, include=FALSE}
# Trying again with GAM

# Check to see if we are using the right formula
# Example to visualize the data
ggplot(summary_data, aes(x = Distance_from_PFU_cm, y = mean_VPFU)) +
  geom_smooth(method = "lm", size = 1)


library(ggplot2)
library(mgcv) # For GAM smoothing

ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  # Shaded area representing ± 1 standard deviation
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = TRUE) +
  # Solid line for the mean intensity
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), method = "gam", formula = y ~ s(x, bs = "cs", k = 3), linewidth = 1) +
  geom_smooth(aes(y = lower_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs", k = 3), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs", k = 3), linetype = "dashed") +
  # Customizing colors to match the Magalie aesthetic
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "darkgreen")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  # Axis labels and title
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  # Theme adjustments
  theme_minimal(base_size = 14) +
  theme(legend.title = element_blank(), legend.position = "top", plot.title = element_text(size = 14, face = "bold", hjust = -0.1))
```



```{r setup, include=TRUE}

library(ggplot2)
library(mgcv) # For GAM smoothing
ggplot(summary_data, aes(x = Distance_from_PFU_cm)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = "Standard deviation"), alpha = 0.2, show.legend = TRUE) +
  geom_smooth(aes(y = mean_VPFU, color = "Mean"), method = "gam", formula = y ~ s(x, bs = "cs", k = 3), linewidth = 1) 
geom_smooth(aes(y = lower_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs", k = 3), linetype = "dashed") +
  geom_smooth(aes(y = upper_bound, color = "Standard deviation"), method = "gam", formula = y ~ s(x, bs = "cs", k = 3), linetype = "dashed") +
  scale_color_manual(values = c("Mean" = "green", "Standard deviation" = "darkgreen")) +
  scale_fill_manual(values = c("Standard deviation" = "lightgreen")) +
  labs(x = "Distance from PFU center (cm)", y = "Virospores") +
  theme_minimal(base_size = 14) +
  theme(legend.title = element_blank(), legend.position = "top", plot.title = element_text(size = 14, face = "bold", hjust = -0.1))
