---
title: "Virospore annulus multi-quant for Magalie"
author: "Joy O'Brien"
date: "2025-03-22"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/GitHub/VirosporeQuant/Magalie_virospore_quant")
```

```{r}
# Load Library
library(readxl)
library(dplyr)
library(tidyr)  
library(vegan)
library(ggplot2)
library(mgcv)

# Read in data
viro_quant <- read_excel("~/GitHub/VirosporeQuant/Magalie_virospore_quant/Data/Annulus_Quant/20250317_virospore_multiquant_magalie.xlsx", sheet = "raw_data_joy")
```

Troubleshooting the back-calculation process
- Because the pipette tip with the sample was dispensed into 1.5 mL of buffer, we need to account for that as the stock solution
- The total dilution factor is from the original stock solution to the plated sample

ORDER OF OPERATIONS FOR CALCULATIONS
- CALCULATE CFU/mL and PFU/mL FOR EACH DILUTION
- CALCULATE THE AMOUNT OF CFU/mL or PFU/mL IN THE STOCK SOLUTION (1.5 ML)
- Calculate the amount of CFU or PFU in the cm^2 sampling area

1. Obtain CFU and PFU measurements for each sample
1A. Calculate CFU/PFU per mL for each dilution
```{r}
# Calculate CFU and PFU of the original suspension which is # of PFU or CFU / vol plated x total dilution factor
viro_quant$Free_PFU_mL_3rep <- viro_quant$Free_PFU_10e3 / (0.1 * (10^-3))
#viro_quant$Free_PFU_mL_5rep <- viro_quant$Free_PFU_10e5 / (0.1 * (10^-5))
#viro_quant$Free_PFU_mL_7rep <- viro_quant$Free_PFU_10e7 / (0.1 * (10^-7))

# Uninfected spores
viro_quant$Spores_CFU_mL_3rep <- viro_quant$Spores_CFU_10e3 / (0.25 * (10^-3))
#viro_quant$Spores_CFU_mL_5rep <- viro_quant$Spores_CFU_10e5 / (0.25 * (10^-5))
#viro_quant$Spores_CFU_mL_7rep <- viro_quant$Spores_CFU_10e7 / (0.25 * (10^-7))


```

1B. Calculate CFU/mL and PFU/mL per stock solution for each sample
```{r}
viro_quant$Free_PFU_3stock <- viro_quant$Free_PFU_mL_3rep * 1.5
#viro_quant$Free_PFU_5stock <- viro_quant$Free_PFU_mL_5rep * 1.5
#viro_quant$Free_PFU_7stock <- viro_quant$Free_PFU_mL_7rep * 1.5

viro_quant$Spores_CFU_3stock <- viro_quant$Spores_CFU_mL_3rep * 1.5
#viro_quant$Spores_CFU_5stock <- viro_quant$Spores_CFU_mL_5rep * 1.5
#viro_quant$Spores_CFU_7stock <- viro_quant$Spores_CFU_mL_7rep * 1.5
```

1C. Calculate the amount of Free PFU and Spores CFU per cm^2 area
```{r}
# Calculate per area for free phage
viro_quant$Free_PFU_3_cm2 <- viro_quant$Free_PFU_3stock / (0.00785)
#viro_quant$Free_PFU_5_cm2 <- viro_quant$Free_PFU_5stock / (0.00785)
#viro_quant$Free_PFU_7_cm2 <- viro_quant$Free_PFU_7stock / (0.00785)

# For free spores
viro_quant$Spores_CFU_3_cm2 <- viro_quant$Spores_CFU_3stock / (0.00785)
#viro_quant$Spores_CFU_5_cm2 <- viro_quant$Spores_CFU_5stock / (0.00785)
#viro_quant$Spores_CFU_7_cm2 <- viro_quant$Spores_CFU_7stock / (0.00785)
```


1D. Calculations for undiluted samples (virospore samples)
```{r}
# Since the dilution factor is 1 (10^0) we just really need to account for the volume plated
viro_quant$VPFU <- viro_quant$VPFU / (0.25) # accounting for volume plated 250 ul 

# And then calculate for the final amount of vpfu in the stock (1.5ml)
viro_quant$VPFU_stock <- viro_quant$VPFU * 1.5

# And then finally take into account the area sampled 
viro_quant$VPFU_cm2 <- viro_quant$VPFU / (0.00785)
```

```{r}
# Plotting Joy's counts
viro_quant.new <- viro_quant[, c("Plaque", "Sample", "Region", "VPFU_cm2", "Distance_from_PFU_center_cm", "Free_PFU_3_cm2", "Spores_CFU_3_cm2")]

# Calculate total spores with VPFU and the averages 
viro_quant.new$Spores_total <- viro_quant.new$Spores_CFU_3_cm2 + viro_quant.new$VPFU_cm2
```


2. Prepare to plot the data 
2A. Create new data frame and average the counts from the technical replicates (diluted samples)
```{r}
# New data frame
viro_quant.new <- viro_quant[, c("Plaque", "Sample", "Region", "VPFU_cm2", "Distance_from_PFU_center_cm", "Free_PFU_3_cm2","Free_PFU_5_cm2", "Free_PFU_7_cm2", "Spores_CFU_3_cm2", "Spores_CFU_5_cm2", "Spores_CFU_7_cm2")] #"total_spores_3rep", #"total_spores_5rep", #"total_spores_7rep

# Average the diluted replicate samples
viro_quant.new$Free_PFU_Avg <- rowMeans(viro_quant.new[, c("Free_PFU_3_cm2", "Free_PFU_5_cm2", "Free_PFU_7_cm2")], na.rm = TRUE)
viro_quant.new$Spores_CFU_Avg <- rowMeans(viro_quant.new[, c("Spores_CFU_3_cm2", "Spores_CFU_5_cm2", "Spores_CFU_7_cm2")], na.rm = TRUE)

# Calculate total spores with VPFU and the averages 
viro_quant.new$Spores_total <- viro_quant.new$Spores_CFU_3_cm2 + viro_quant.new$VPFU_cm2

# Average the total spores count 
#viro_quant.new$total_spores_Avg <- rowMeans(viro_quant.new[, c("total_spores_3rep", "total_spores_5rep", "total_spores_7rep")], na.rm = TRUE)
```

2B. Plot 
```{r}
# Convert data to long format
viro_quant.long <- viro_quant.new %>%
  pivot_longer(cols = c(VPFU_cm2, Free_PFU_3_cm2, Spores_CFU_3_cm2), 
               names_to = "variable", 
               values_to = "value") #%>%
               #mutate(value = log(value)) # for log transforming data


region_shading <- data.frame(
  Region = c("Center", "Annulus", "Lawn"),
  xmin = c(0, 0.05, 0.1),   # Starting x-value for each region
  xmax = c(0.05, 0.1, 0.2)  # Ending x-value for each region
)
region_shading$Region <- factor(region_shading$Region, levels = c("Center", "Annulus", "Lawn"))

# Plot
ggplot(viro_quant.long, aes(x = Distance_from_PFU_center_cm, y = value, color = variable, group = variable)) +
  # Add shaded regions using geom_rect()
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            alpha = 0.3) +
  # Line and point layers
  geom_line(linewidth = 1.2) +  
  #geom_smooth() +
  geom_point(aes(color = variable), size = 2) + 
  # Custom fill colors for shaded regions
  scale_fill_manual(values = c("Center" = "lightblue", "Annulus" = "lightgreen", "Lawn" = "lightpink")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  # Labels 
  labs(title = "Virospore Multi-Quant Plot Plaque A", 
       x = "Distance from PFU center (cm)", 
       y = "Virospores, Free Phage, Uninfected Spores per cm^2", 
       color = "Measurement") +
  theme_minimal()
```

2C. Plot all samples as individuals
```{r}
viro_quant.long.all <- viro_quant.new %>%
  pivot_longer(cols = c(VPFU_cm2, Free_PFU_3_cm2, Free_PFU_5_cm2, Free_PFU_7_cm2, Free_PFU_Avg, Spores_CFU_3_cm2, Spores_CFU_5_cm2, Spores_CFU_7_cm2, Spores_CFU_Avg),
   names_to = "variable", 
               values_to = "value")
# Plot
ggplot(viro_quant.long.all, aes(x = Distance_from_PFU_center_cm, y = value, color = variable, group = variable)) +
  # Add shaded regions using geom_rect()
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            alpha = 0.3) +
  # Line and point layers
  geom_line(linewidth = 1.2) +  
  geom_point(aes(color = variable), size = 2) + 
  # Custom fill colors for shaded regions
  scale_fill_manual(values = c("Center" = "lightblue", "Annulus" = "lightgreen", "Lawn" = "lightpink")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  # Labels 
  labs(title = "Virospore Multi-Quant Plot Plaque A", 
       x = "Distance from PFU center (cm)", 
       y = "Virospores, Free Phage, Uninfected Spores per cm^2", 
       color = "Measurement") +
  theme_minimal()
```


```{r}

```





