---
title: "Macroplate analysis for Magalie"
author: "Joy O'Brien"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/GitHub/danger.signal/2.square_plate")
```

0. Set plot theme
```{r}
library(paletteer)
library(ggplot2)
theme_set(theme_classic() +
  theme(axis.title = element_text(size = 16),
        axis.title.x = element_text(margin = margin(t = 15, b = 15)),
        axis.title.y = element_text(margin = margin(l = 15, r = 15)),
        axis.text = element_text(size = 13),
        axis.text.x = element_text(margin = margin(t = 5)),
        axis.text.y = element_text(margin = margin(r = 5)),
        #axis.line.x = element_line(linewidth = 1),
        #axis.line.y = element_line(linewidth = 1),
        axis.ticks.y.right = element_line(linewidth = 1),
        axis.text.y.right = element_text(size = 13, margin = margin(l = 5)),
        axis.line.y.right = element_line(linewidth = 0.1),
        axis.ticks.x.top = element_line(linewidth = 1),
        axis.text.x.top = element_text(size = 13, margin = margin(b = 5)),
        axis.line.x.top = element_line(linewidth = 0.1),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_line(linewidth = 1),
        axis.ticks.y = element_line(linewidth = 1),
        axis.ticks.length = unit(.1, "in"),
        panel.border = element_rect(color = "grey20", fill = NA, linewidth = 1.5),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 14),
        strip.background = element_blank()
        )) 
colors <-paletteer_d("ggthemes::Miller_Stone")

# Emmi's theme set code: 
# theme_set(theme_classic() +
#   theme(axis.title = element_text(size = 16),
#         axis.title.x = element_text(margin = margin(t = 15, b = 15)),
#         axis.title.y = element_text(margin = margin(l = 15, r = 15)),
#         axis.text = element_text(size = 13),
#         axis.text.x = element_text(margin = margin(t = 5)),
#         axis.text.y = element_text(margin = margin(r = 5)),
#         #axis.line.x = element_line(linewidth = 1),
#         #axis.line.y = element_line(linewidth = 1),
#         axis.line.x = element_blank(),
#         axis.line.y = element_blank(),
#         axis.ticks.x = element_line(linewidth = 1),
#         axis.ticks.y = element_line(linewidth = 1),
#         axis.ticks.length = unit(.1, "in"),
#         panel.border = element_rect(color = "grey20", fill = NA, linewidth = 1.5),
#         legend.text = element_text(size = 14),
#         strip.text = element_text(size = 14),
#         strip.background = element_blank()
#         ))
```


1. Load in all data
```{r}
library(readxl)
# Read in the measurements
h11 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/11h_redo.csv")
h12 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/12h_redo.csv")
h13 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/13h_redo.csv")
h14 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/14h_redo.csv")
h15 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/15h_redo.csv")
h16 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/16h_redo.csv")
h17 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/17h_redo.csv")
h18 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/18h_redo.csv")
h19 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/19h_redo.csv")
h23 <- read.csv("~/GitHub/danger.signal/2.square_plate/Data/23h_redo.csv")

# Read in the metadata
meta <- read_xlsx("~/GitHub/danger.signal/2.square_plate/Data/square_plate_meta.xlsx", sheet = "Sheet1")
```

1A. Add a Time column in each of the datasets
```{r}
library(dplyr)
h11 <- h11 %>%
  mutate(Time_h = 11)

h12 <- h12 %>%
  mutate(Time_h = 12)

h13 <- h13 %>%
  mutate(Time_h = 13)

h14 <- h14 %>%
  mutate(Time_h = 14)

h15 <- h15 %>%
  mutate(Time_h = 15)

h16 <- h16 %>%
  mutate(Time_h = 16)

h17 <- h17 %>%
  mutate(Time_h = 17)
h18 <- h18 %>%
  mutate(Time_h = 18)

h19 <- h19 %>%
  mutate(Time_h = 19)

h23 <- h23 %>%
  mutate(Time_h = 23)
```

1B. Combine all data to one giant dataset
```{r}
square.plate <- bind_rows(h11,h12,h13,h14,h15,h16,h17,h18,h19,h23)

# Rename X to Sample_ID
square.plate <- square.plate %>%
  rename(Sample_ID = X)

# Remove columns we don't need
#square.plate <- square.plate %>%
 #select(-X, -Time-h)

# Add the metadata 
library(dplyr)

square.plate.meta <- left_join(square.plate, meta, by = c("Sample_ID", "Time_h"))
```

1C. Normalize samples by their dilution factor
```{r}
# Add a column specifying dilution for each sample
library(dplyr)
library(stringr)
library(tidyr)
square.plate.norm <- square.plate.meta %>%
  mutate(
    Dilution = str_extract(Sample, "e\\d+") %>%       # extract 'e1', 'e2', etc.
               str_remove("e") %>%                    # remove the 'e'
               as.numeric() %>%                       # convert to number
               (\(x) 10^(-x))()                       # convert to 10^-x
  )
# Normalize pixel intensity (IntDen) by the dilution factor
square.plate.norm <- square.plate.norm %>%
  mutate(
    Mean_norm = Mean / Dilution,
    IntDen_norm = IntDen / Dilution,
    RawIntDen_norm = RawIntDen / Dilution
  )


# Average the background values for each timepoint and subtract to obtain pixel value
# Compute background averages per timepoint and subtract samples based on the mean background for each timepoint
background_means <- square.plate.norm %>%
  filter(Treatment == "Background") %>%
  group_by(Time_h) %>%
  summarise(across(c(Area, Mean, Min, Max, IntDen, RawIntDen), mean, .names = "bg_{.col}"))

# Merge background values back into original dataset by Time_h
square.plate.bg <- square.plate.norm %>%
  left_join(background_means, by = "Time_h") %>%
  mutate(
    Area_corr = Area - bg_Area,
    Mean_corr = Mean - bg_Mean,
    Min_corr = Min - bg_Min,
    Max_corr = Max - bg_Max,
    IntDen_corr = IntDen - bg_IntDen,
    RawIntDen_corr = RawIntDen - bg_RawIntDen
  )

# Create dataset with only the corrected values and sample identifiers
square.plate.corr <- square.plate.bg %>% # for some reason this code is broken so we have to force dplyr select
  select(Sample_ID, Sample, Treatment, Time_h,
          Area_corr, Mean_corr, Min_corr, Max_corr,
          IntDen_corr, RawIntDen_corr)

# square.plate.corr <- square.plate.bg %>%
#   dplyr::select(Sample_ID, Sample, Treatment, Time_h,
#                 Area_corr, Mean_corr, Min_corr, Max_corr,
#                 IntDen_corr, RawIntDen_corr)

```


```{r}
# Rescale?
#rescale_minmax <- function(x, lower = 1, upper = 10) {
  #return(lower + (x - min(x)) / (max(x) - min(x)) * (upper - lower))
#}
# Trying Z-score normalization/rescaling
#zscore_normalization <- function(x) {
 # return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
#}

# Specify columns to scale
#columns_to_scale <- c("IntDen")
#square.plate.meta[columns_to_scale] <- lapply(square.plate.meta[columns_to_scale], rescale_minmax)

```

1E. Plotting/Edits for publication worthy figure
```{r}
library(dplyr)
library(tidyr)

sp.summary <- square.plate.corr %>%
  group_by(Treatment, Time_h) %>%
  summarise(
    mean_IntDen = mean(IntDen_corr, na.rm = TRUE),
    sem_IntDen = sd(IntDen_corr, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
```

1F. Making more informed treatment labels (renaming the treatments)
```{r}

library(dplyr)

sp.summary <- sp.summary %>%
  mutate(Treatment = recode(Treatment,
                            "Background" = "Background",
                            "DSM" = "Fresh media",
                            "Non_reporter_control" = "Non-GFP",
                            "Phage_Spent_DSM" = "Phage lysate",
                            "Scells_Spent_DSM" = "Dead cells",
                            "Spent_DSM" = "Spent media",
                            "water" = "Water"))
```


1G.Plot for publication
```{r}
sp.time.plot <- ggplot(sp.summary, aes(x = Time_h, y = mean_IntDen, color = Treatment)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = mean_IntDen - sem_IntDen, ymax = mean_IntDen + sem_IntDen),
                width = 0.2, color = "black") +
  labs(
    title = "",
    x = "Time (h)",
    y = "GFP (a.u.)",
    color = "Treatment"
  ) +
  scale_x_continuous(
    limits = c(10,28),
    breaks = c(10, 15, 20, 25),
    sec.axis = dup_axis(name = NULL, labels = NULL)
  ) +
  scale_color_manual(values = colors) + # This is the only way I know how to get the tick marks (How does Emmi do it)
  scale_y_continuous(sec.axis = dup_axis(name = NULL, labels = NULL)) +
  #theme()
  #axis.line.y.right = element_line(linewidth = 0.1), 
  #axis.line.x.top = element_line(linewidth = 0.1)
  theme(legend.position = "none") +
  annotate("text", x = 26, y = 8900, label = "Dead cells", color = "#BFBB60FF", size = 3) +
  annotate("text", x = 26, y = 8500, label = "Spent media", color =  "#F47942FF", size = 3) +
    annotate("text", x = 26, y = 8100, label = "Phage lysate", color =  "#638B66FF", size = 3) +
  annotate("text", x = 26, y = 7700, label = "Fresh media", color = "#849DB1FF", size = 3) +
  annotate("text", x = 26, y = 6500, label = "Water", color = "#FBB04EFF", size = 3) +
  annotate("text", x = 26, y = 900, label = "Non-GFP", color = "#A2CEAAFF", size = 3) + 
  annotate("text", x = 26, y = 0, label = "Background", color = "#4F6980FF", size = 3)
  
sp.time.plot  
#+
  #annotate("text", x = 20, y = 10000, label = "Rm-ANOVA p = 0.0001*", color = "black", size = 5)


```

Extra: Plot (old)
```{r}
library(ggplot2)

ggplot(square.plate.corr, aes(x = Time, y = IntDen_corr, color = Treatment)) +
  geom_line() +
  #geom_smooth(span = 0.4, alpha = 0.2) +
  #scale_y_log10() +
  #geom_smooth(span, se = TRUE, alpha = 0.2) +
  geom_point() +
  geom_errorbar(aes(ymin = IntDen_corr - SEM, ymax = IntDen_corr + SEM),
              width = 0.005, color = "black") +
  labs(
    title = "",
    x = "Time (hours)",
    y = "GFP Intensity",
    color = "Treatment"
  ) +
    theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    panel.grid.major = element_blank(),  # removes major grid lines
    panel.grid.minor = element_blank(),  # removes minor grid lines
    panel.background = element_blank(),  # removes the background color
    plot.background = element_blank(),  # removes background around the plot
    theme_minimal()  # minimal theme
  ) + 
   annotate("text", x = 20, y = 10000, label = "Rm-ANOVA p = 0.0001*", color = "black", size = 5)
```


2. Prep for RM-ANOVA
```{r}
# Before running repeated measures anova remove the WT control from the samples (same thing as a blank?)
square.plate.meta.noWT <- square.plate.corr %>%
  filter(Treatment != "Non_reporter_control")
# And remove the blank 
square.plate.meta.noWTBL <- square.plate.meta.noWT %>%
  filter(Treatment != "Background")

square.plate.meta.noWTBL <- square.plate.meta.noWTBL %>%
  filter(Treatment != "water")

```

2A. Repeated-Measures ANOVA
```{r}
aov_result_sp <- aov(IntDen_corr ~ Treatment * Time_h + Error(Sample / Time_h), data = square.plate.meta.noWTBL)
summary(aov_result_sp)

library(emmeans)

emmeans_result_sp <- emmeans(aov_result_sp, ~ Treatment | Time_h)
pairs(emmeans_result_sp)
```

3. Set up to apply data to Gompertz (running into issues here)
```{r}
setwd("~/GitHub/danger.signal/2.square_plate/GrowthCurves")
```

```{r}
# Load packages and functions
require("png")
require("dplyr")
require("grid")
require("gtools")
require("nlme")
require("MuMIn")
require("bbmle")
source("~/GitHub/danger.signal/2.square_plate/GrowthCurves/bin/modified_Gomp_diagnostic3.R")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) 100*( sd(x)/mean(x))
```

3A. Load data (already loaded)
```{r}
square.plate.corr
```

3B. Wrangle data (group samples by treatment)
```{r}
library(dplyr)
# Trying this out first with water data first to make sure it goes smoothly
sp.water <- square.plate.corr %>%
  filter(Treatment %in% c("water"))
sp.DSM <- square.plate.corr %>%
  filter(Treatment %in% c("DSM"))
sp.spent.DSM <- square.plate.corr %>%
  filter(Treatment %in% c("Spent_DSM"))
sp.SC.spent <- square.plate.corr %>%
  filter(Treatment %in% c("Scells_Spent_DSM"))
sp.phage <- square.plate.corr %>%
  filter(Treatment %in% c("Phage_Spent_DSM"))

# Get rid of unnecessary data
sp.water.clean <- sp.water %>%
  select(Time_h, Sample, IntDen_corr)

sp.DSM.clean <- sp.DSM %>%
  select(Time_h, Sample, IntDen_corr)

sp.spent.DSM.clean <- sp.spent.DSM %>%
  select(Time_h, Sample, IntDen_corr)

sp.SC.spent.clean <- sp.SC.spent %>%
  select(Time_h, Sample, IntDen_corr)

sp.phage.clean <- sp.phage %>%
  select(Time_h, Sample, IntDen_corr)
``` 

3B. Continue to wrangle data (wide format)
```{r}

# Transform to wide format so we can run the model
library(tidyr)
sp.water.wide <- sp.water.clean %>%
  pivot_wider(names_from = Sample, values_from = IntDen_corr)

sp.DSM.wide <- sp.DSM.clean %>%
  pivot_wider(names_from= Sample, values_from = IntDen_corr)

sp.spent.wide <- sp.spent.DSM.clean %>%
  pivot_wider(names_from= Sample, values_from = IntDen_corr)

sp.SC.spent.wide <- sp.SC.spent.clean %>%
  pivot_wider(names_from= Sample, values_from = IntDen_corr)

sp.phage.wide <- sp.phage.clean %>%
  pivot_wider(names_from= Sample, values_from = IntDen_corr)

# Change the Time_h to Time
sp.water.wide <- sp.water.wide %>%
  rename(Time = Time_h)

sp.DSM.wide <- sp.DSM.wide %>%
  rename(Time = Time_h)

sp.spent.wide <- sp.spent.wide %>%
  rename(Time = Time_h)

sp.SC.spent.wide <- sp.SC.spent.wide %>%
  rename(Time = Time_h)

sp.phage.wide <- sp.phage.wide %>%
  rename(Time = Time_h)

# OKAY THIS IS THE FORMAT WE WANT!
```

3C. Run Gompertz model (under construction)
```{r}
# Running Gompertz on water data to make sure it runs smoothly
setwd("~/GitHub/danger.signal/2.square_plate/GrowthCurves/bin")
sp.water.1 <- growth.modGomp(input = sp.water.wide, output.name = "sp.water.1.parms",
                 synergy = F, temp = F, smooth = T, trim = T)

# Same error as the flow cytometry data -- > go under Gompertz hood
[1] Starting Sample Water_e1 (1 of 6)
Error in ts(x) : 'ts' object must have one or more observations

# Checking some things that may be associated with the error
samples[!samples %in% colnames(sp.water.wide)] # okay samples are not found so this means
samples <- setdiff(colnames(sp.water.wide), "Time")
samples[!samples %in% colnames(sp.water.wide)]

#sp.water.wide$Time<- format(strptime(sp.water.wide$Time, format = "%H"), "%H:%M")
#head(sp.water.wide, header = T)

sp.water.wide.df <- as.data.frame(sp.water.wide)
# Okay we specified samples, now let's try again
setwd("~/GitHub/danger.signal/2.square_plate/GrowthCurves/bin")
sp.water.1 <- growth.modGomp(input = sp.water.wide.df, output.name = "sp.water.1.parms",
                 synergy = F, temp = F, smooth = F, trim = T)


# Okay things that I needed to fix: 
# as.data.frame needed for running gompertz
# specified samples
# Smoothing = F

# Further debugging
#Error in ts(x) : 'ts' object must have one or more observations

print(head(sp.water.wide))  # Inspect the first few rows
print(length(sp.water.wide$y))  # Check the number of observations
print(sum(is.na(sp.water.wide$y)))  # Check for missing values
```


4. Obtain parameters via GrowthCurver package (back up method)
```{r}
# Make all data.frames
sp.water.wide.df <- as.data.frame(sp.water.wide)
sp.DSM.wide.df <- as.data.frame(sp.DSM.wide)
sp.spent.wide.df <- as.data.frame(sp.spent.wide)
sp.SC.spent.wide.df <- as.data.frame(sp.SC.spent.wide)
sp.phage.wide.df <- as.data.frame(sp.phage.wide)


library(growthcurver)
# Time should not be in 00:00 format

water.plate <- sp.water.wide.df
  names(water.plate)
  
DSM.plate <- sp.DSM.wide.df
  names(DSM.plate)
  
spent.plate <- sp.spent.wide.df 
  names(spent.plate)

SC.spent.plate <- sp.SC.spent.wide.df
  names(SC.spent.plate)

phage.plate <- sp.phage.wide.df
  names(phage.plate)
  
#obtain growth parms

water.plate.summary <-SummarizeGrowthByPlate(plate = water.plate)
head(water.plate.summary)

DSM.plate.summary <- SummarizeGrowthByPlate(plate = DSM.plate)
head(DSM.plate.summary)

spent.plate.summary <- SummarizeGrowthByPlate(plate = spent.plate)
head(spent.plate.summary)

SC.spent.plate.summary <- SummarizeGrowthByPlate(plate = SC.spent.plate)
head(SC.spent.plate.summary)

phage.plate.summary <- SummarizeGrowthByPlate(plate = phage.plate)
head(phage.plate.summary)
```

4A. Calculate lag time (onset time)
```{r}
# Calculate lag time 
water.plate.summary$lag_time <- water.plate.summary$t_mid - (1 / water.plate.summary$r)
DSM.plate.summary$lag_time <- DSM.plate.summary$t_mid - (1 / DSM.plate.summary$r)
spent.plate.summary$lag_time <- spent.plate.summary$t_mid - (1 / spent.plate.summary$r)
SC.spent.plate.summary$lag_time <- SC.spent.plate.summary$t_mid - (1 / SC.spent.plate.summary$r)
phage.plate.summary$lag_time <- phage.plate.summary$t_mid - (1 / phage.plate.summary$r)

# Combine into one df but first add treatments
water.plate.summary$Treatment <- "Water"
DSM.plate.summary$Treatment <- "DSM"
spent.plate.summary$Treatment <- "Spent"
SC.spent.plate.summary$Treatment <- "SC_Spent"
phage.plate.summary$Treatment <- "Phage"

lag_data <- bind_rows(
  water.plate.summary,
  DSM.plate.summary,
  spent.plate.summary,
  SC.spent.plate.summary,
  phage.plate.summary
)
```

Extra: Quick look at the data
```{r}
library(ggplot2)

ggplot(lag_data, aes(x = Treatment, y = lag_time, fill = Treatment)) +
  geom_boxplot() +
  labs(title = "",
       x = "Treatment",
       y = "Lag Time (h)") +
  theme_minimal() +
  theme(legend.position = "none")

```

5. ANOVA onset time
```{r}
# ANOVA
anova.lag <- aov(lag_time ~ Treatment, data = lag_data) # ANOVA
summary(anova.lag)
tukey_result.sp <- TukeyHSD(anova.lag) # Obtain tukey result

# Obtain group numbers for significance
library(emmeans)
library(multcomp)

# Get estimated marginal means
emm <- emmeans(anova.lag, pairwise ~ Treatment)

# Compact Letter Display (CLD)
# NOTE: adjust = "tukey" ensures same correction as TukeyHSD
cld_result <- cld(emm$emmeans, adjust = "tukey")

# View results
print(cld_result)


```

4B. Renaming treatments for publication
```{r}

library(dplyr)

lag_data <- lag_data %>%
  mutate(Treatment = recode(Treatment,
                            "Background" = "Background",
                            "DSM" = "Fresh media",
                            "Non_reporter_control" = "Non-GFP",
                            "Phage" = "Phage lysate",
                            "SC_Spent" = "Dead cells",
                            "Spent" = "Spent media",
                            "water" = "Water"))
```


4C. Reordering appearance of treatments and plotting (COME BACK TO THIS NEED TO ADD TOP AXIS TICK MARKS)
```{r}
lag_data$Treatment <- factor(lag_data$Treatment, levels =c("Water", "Fresh media", "Spent media", "Dead cells", "Phage lysate"))
library(legendry)
# Okay the above code adds the top ticks but does not preserve the bottom x axis 
sp.box.plot <- ggplot(lag_data, aes(x = Treatment, y = lag_time)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  scale_x_discrete(labels = c(
    "Water",
    "Fresh\nmedia",
    "Spent\nmedia",
    "Dead\ncells",
    "Phage\nlysate"
  )) +
  labs(
    x = "Treatment",
    y = "Sporulation onset (h)",
    title = ""
  ) +
  scale_y_continuous(sec.axis = sec_axis(~., labels = NULL)) +
  annotate("text", x = 4, y = 18, label = "ANOVA p-value = 0.1", color = "black") +
  theme(
    axis.ticks.x.top = element_line(),       # Show top ticks
    axis.line.x.top = element_line(),        # Draw top axis line
    axis.text.x.top = element_blank(),       # Optional: remove top labels
    axis.title.x.top = element_blank()       # Optional: remove top title
  ) +
  guides(x.sec = legendry::guide_axis_base(key = key_map(data = lag_data, aesthetic = unique(Treatment), label = c("", "", "", "", ""))))
print(sp.box.plot)
```


5. Plotting the two plots together (square plate time series plot and box plot) + including photo of experiment
```{r}
library(ggplot2)
library(cowplot)
library(magick)
plate.img <- image_read("~/GitHub/danger.signal/2.square_plate/Data/23_h-1.tif")

#Convert image to cowplot object
# Step 2: Convert image to a cowplot object
img_plot <- ggdraw() + draw_image(plate.img)
empty.plot <- ggdraw()
# Okay actually we will add the image later, create an empty plot 
# Step 3: Combine all plots with the image on the left
combined_plots <- ggdraw() +
  draw_plot(empty.plot, x = 0,   y = 0, width = 1/3, height = 1) +  # Left image
  draw_plot(sp.time.plot,  x = 1/3, y = 0, width = 1/3, height = 1) +  # Center plot
  draw_plot(sp.box.plot,   x = 2/3, y = 0, width = 1/3, height = 1) +  # Right plot
  theme(plot.background = element_rect(fill= "white", color = NA)) +
  draw_plot_label(label = c("(A)", "(B)", "(C)"), size = 16,
                  x = c(0.05, 0.38, 0.71), y = c(1, 1, 1))

print(combined_plots)

ggsave("~/GitHub/danger.signal/2.square_plate/Figures/combined_plotswspace.png", combined_plots, width = 12, height = 4)
```

Extra: Original version of the above code
```{r}
# Need to adjust to add the picture
#dev.new(width=5, height=4)
dev.new(width=5,height=4,noRStudioGD = TRUE)

# Include a space before the two plots for the picture
# Use ggdraw and draw_plot
combined_plots <- ggdraw() +
  draw_plot(sp.time.plot, x = 0,   y = 0, width = 0.5, height = 1) +  # Left plot
  draw_plot(sp.box.plot, x = 0.5, y = 0, width = 0.5, height = 1) +    # Right plot
  theme(plot.background = element_rect(fill= "white", color = NA)) +
  #draw_plot_label(label = c("(A)", "(B)", "(C)"), size = 20, x=c(0.07, 0.53), y = c(1, 1))

# Show the combined plot
print(combined_plots)

#ggsave("~/GitHub/danger.signal/2.square_plate/Figures/combined_plots.png", combined_plots, width = 12, height = 4)

```

