---
title: "Sporulation curve on agar surface"
output:
  html_document:
    df_print: paged
  html_notebook:
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

```{r, , include=FALSE}
library(tidyverse, quietly = TRUE)
library(cowplot, quietly = TRUE)
library(outliers, quietly = TRUE)
library(here, quietly = TRUE)
```


```{r consolidate data, echo=FALSE, message=FALSE}
# get all the file names to collect
files <- list.files(here("data/output/"), pattern = ".csv",full.names = T, recursive = T)
d <- data.frame()

#This loop collects all the data into a single dataframe

for (f in files){
   tmp <- read_csv(f)


   #add experimental time (hr) as number
     tmp <- tmp %>%
     mutate(time.hr = parse_number(t.exp)) %>% 
       relocate(time.hr, .after = "t.exp")

   d <- rbind(d, tmp)

      }
rm(tmp)

# #check completeness of data
# d%>%
#    group_by(exp.plate, time.hr, bio.rep)%>%
#    summarise(n=n()) %>% 
#   view()
# 
# d%>%
#    group_by(exp.plate, time.hr)%>%
#    filter(bio.rep != "blank") %>% 
#    summarise(n=n()) %>% 
#   view()

```

clean up data

```{r}

d.clean <-
   d%>%
   # make spore NAs into 1 (0 on log scale)
   mutate(spore = if_else(!is.na(spore), 1, spore))%>%
   # make spore counts lower than 100 events 1 (0 on log scale)
   mutate(spore = if_else(spore<100, 1, spore) )%>%
   # make veg counts lower than 100 events 1 (0 on log scale)
   mutate(veg = if_else(veg<100, 1, veg) ) %>% 
  # unique sample name
  mutate(unq.sample = str_c(t.exp, bio.rep, exp.plate, sep = "_")) %>% 
  # blanks are not useful for this analysis
  filter(bio.rep!="blank") %>% 
  
  mutate(spore = if_else(is.na(spore), 0.1, spore),
         spore.ml = if_else(is.na(spore.ml), 0.1, spore.ml)) %>% 
  
  #error in exp.plate assginment
  mutate(exp.plate = str_replace(exp.plate, "plt23", "plt3"))

```

Look at data

```{r}   
   

d.clean%>%
   pivot_longer(cols = c(veg.ml, spore.ml),
                names_to = "pop", values_to = "cell.ml" )%>%
  # mutate()
   ggplot(aes(x=bio.rep, y=cell.ml))+
   geom_point(aes(color=pop, shape=rep))+
   # facet_wrap(~interaction(time.hr,exp.plate))+
  facet_wrap(~time.hr+exp.plate)+
   theme_cowplot()+
   panel_border()+
   scale_y_log10()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
         strip.background = element_blank(),
         strip.text = element_text(size=8, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")))
```
Mostly technical replicates are similar (=rep). Some are off but that should mostly be solved by removing outliers.
Biological replication also look good.

# N samples
```{r}
d.clean%>%
  group_by(time.hr, exp.plate) %>% 
  summarise(n=n()) %>% 
  pivot_wider(names_from = exp.plate,values_from = n) %>% view
  ggplot(aes(time.hr, volume.nL/1000))+
  geom_point(aes(color=dilution))+
  facet_wrap(~exp.plate, ncol = 2)+
  theme_bw()
```

dilutions
```{r}
d.clean%>%
  ggplot(aes(time.hr, volume.nL/1000))+
  # ggplot(aes(time.hr, total.events/1000))+
  geom_point(aes(shape = exp.plate,color=dilution))+
  # facet_wrap(~exp.plate, ncol = 2)+
  theme_bw()
```


# Replication outliers
I have made 3 technical replicate measurements in the flow-cytometer for each sample. There are some measurements which are obviously way off, e.g. in "WED12" samples some wells did not get SYBR to to pipetting error. To get rid of these I will choose from each triplicate the 2 points that are in best agreement and remove the 3rd point. There is a function to idenify such a point in pacakge 'outliers'
> outlier {outliers} Finds value with largest difference between it and sample mean, which can be an outlier.

> logical: if set to TRUE, gives vector of logical values, and possible outlier position is marked by TRUE

I will apply the outlier filtering on the number of spores (spore/mL)

```{r id outliers, echo=FALSE}
d2 <- 
  d.clean %>% 
  group_by(unq.sample) %>% 
  filter (sybr.modes == median(sybr.modes)) 


d3 <- tibble()
for (i in unique(d2$unq.sample)){
  
  tmp <- d2 %>% 
    filter (unq.sample ==i)
  

  if (nrow(tmp)<3){
    tmp$out.lier <- FALSE
  } else if (sum(tmp$spore.ml)<300){
    tmp <- tmp %>% 
    mutate(out.lier = outlier(veg.ml, logical = T)) 
  } else {
    tmp <- tmp %>% 
    mutate(out.lier = outlier(spore.ml, logical = T)) 
  }

   d3 <- bind_rows(d3,tmp)
    
}



d3%>%
   ggplot(aes(x=bio.rep, y=veg.ml))+
    geom_jitter(aes(color=out.lier), shape=21, width = 0.2, height = 0)+
  facet_wrap(~time.hr+exp.plate)+
   theme_cowplot()+
   panel_border()+
   scale_y_log10()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
         strip.background = element_blank(),
         strip.text = element_text(size=8, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")))

d3%>%
   ggplot(aes(x=bio.rep, y=spore.ml))+
   geom_jitter(aes(color=out.lier), shape=21, width = 0.2, height = 0)+
  facet_wrap(~time.hr+exp.plate)+
   theme_cowplot()+
   panel_border()+
   scale_y_log10()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
         strip.background = element_blank(),
         strip.text = element_text(size=8, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")))
```


```{r}

d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml),names_to = "pop", values_to = "n.ml") %>% 
  
    group_by(time.hr, pop) %>%
  summarise(m = mean(n.ml), v= sd(n.ml) ) %>% 

  
  ggplot(aes(time.hr, m, color = pop))+
  geom_pointrange(aes(ymin = m-v, ymax = m+v))+
  stat_smooth()+
  # scale_y_log10()+

  theme_cowplot()+
  panel_border(color = "black")
  # facet_wrap(~bio.rep)+
  # theme(strip.background = element_blank())
  

  
```
```{r}

d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml) %>% round(), spore.ml = mean(spore.ml)%>% round()) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml, 
          perc.spore =spore.ml*100/cell.ml ) %>% 
  
  # write_csv(here("data","Sporulation_by_Time_FCM.csv"))

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml, perc.spore),
               names_to = "pop", values_to = "n.ml") %>% 
  
  # mutate(n.ml = if_else(pop == "perc.spore", n.ml, log10(n.ml))) %>% 
  group_by(pop) %>% 
    ggplot(aes(time.hr, n.ml))+
  geom_point(aes(color = exp.plate))+
  stat_smooth()+
  facet_wrap(~pop, scales = "free_y")+
  theme_cowplot()+
  panel_border(color = "black")+
   theme(strip.background = element_blank())
  
```

```{r}

d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml, 
          perc.spore =spore.ml*100/cell.ml ) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml, perc.spore),
               names_to = "pop", values_to = "n.ml") %>% 
  group_by(pop) %>% 
  mutate(rel.abundance = n.ml/max(n.ml)) %>% 
    ggplot(aes(time.hr, rel.abundance))+
  geom_point(aes(color = exp.plate))+
  # stat_smooth()+
    stat_smooth(formula = "y ~ x", method = "glm", 
              method.args = list(family="quasibinomial"), se = T) +
  facet_wrap(~pop, scales = "free_y")+

  theme_cowplot()+
  panel_border(color = "black")+
   theme(strip.background = element_blank())
  
```

```{r}
library(gcplyr)


```

