---
title: "Sporulation curve on agar surface"
output:
  html_document:
    df_print: paged
  html_notebook:
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

```{r, , include=FALSE}
library(tidyverse, quietly = TRUE)
library(cowplot, quietly = TRUE)
library(outliers, quietly = TRUE)
library(here, quietly = TRUE)
```


```{r consolidate data, echo=FALSE, message=FALSE}
# get all the file names to collect
files <- list.files(here("data/output/"), pattern = ".csv",full.names = T, recursive = T)
d <- data.frame()

#This loop collects all the data into a single dataframe

for (f in files){
   tmp <- read_csv(f)


   #add experimental time (hr) as number
     tmp <- tmp %>%
     mutate(time.hr = parse_number(t.exp)) %>% 
       relocate(time.hr, .after = "t.exp")

   d <- rbind(d, tmp)

      }
rm(tmp)

# #check completeness of data
# d%>%
#    group_by(exp.plate, time.hr, bio.rep)%>%
#    summarise(n=n()) %>%
#   view()
# 
# d%>%
#    group_by(exp.plate, time.hr)%>%
#    filter(bio.rep != "blank") %>% 
#    summarise(n=n()) %>% 
#   view()

```


### Fix a couple of errors

In manual inspection of gate plots ("scatterSYBR" plots) I recognized 2 errors:  

1. Samples swapped. In sample taken on *TUE21* samples of *DSM-br2* were swapped between T24 (plt1) and T12 (plt2). I will resove this by swapping the labels.  

2. T0 samples were not handled well by the classification cluster model (veg vs. spores). It appears that there is misclassification of veg events as spores, while there don't seem to be any spores. I will reassign all the spore counts in these samples to veg.  

3. Pipetting error. In *T12* sample of *br3* in *plate 1* only one of the treatments was initially processed. I caught the error in time and analyzed all the treatments labeling them *br3r*. 

```{r fix1}
d.swapped <- d %>% 
  filter(sample.date == "TUE21" &
           trt == "DSM" &
           bio.rep == "br2") %>% 
  mutate(exp.plate = if_else(well %in% c("A8","B8","C8"), "plt2", "plt1"),
         t.exp = if_else(well %in% c("A8","B8","C8"), "T12", "T24"),
         time.hr = if_else(well %in% c("A8","B8","C8"), 12, 24))

d <- d %>% 
  filter(sample.date != "TUE21" |
           trt != "DSM" |
           bio.rep != "br2") %>% 
  rbind(., d.swapped)

rm(d.swapped)
```

```{r fix2}
d.T0 <- d %>% 
  filter(t.exp == "T00") %>% 
  mutate(veg = veg + spore, 
         veg.ml = veg.ml + spore.ml) %>% 
  mutate(spore = 0, spore.ml = 0)

d <- d %>% 
  filter(t.exp != "T00") %>% 
  rbind(., d.T0)

rm(d.T0)

```

```{r fix3}
d.br3r <- d %>% 
  filter(sample.date == "TUE09" &
          t.exp == "T12" &
          exp.plate == "plt1" &
          str_detect(bio.rep, "br3")) %>% 
  filter(bio.rep == "br3r") %>% 
  mutate(bio.rep = "br3")

d <- d %>% 
  filter(sample.date != "TUE09" |
          t.exp != "T12" |
          exp.plate != "plt1" |
          !str_detect(bio.rep, "br3")) %>% 
  bind_rows(., d.br3r)

rm(d.br3r)
  
```


### clean up data

```{r}

d.clean <-
   d%>%
   # make NAs into 1 (0 on log scale)
  mutate(spore = if_else(is.na(spore), 1, spore))%>%
  mutate(veg = if_else(is.na(veg), 1, veg))%>%
   # make spore counts lower than 1000 events 1 (0 on log scale)
   mutate(spore = if_else(spore<1000, 1, spore) )%>%
  mutate(spore.ml = if_else(spore<1000, 1, spore.ml) )%>%
   # make veg counts lower than 1000 events 1 (0 on log scale)
   mutate(veg = if_else(veg<1000, 1, veg) ) %>% 
  mutate(veg.ml = if_else(veg<1000, 1, veg.ml) ) %>% 
  # unique sample name
  mutate(unq.sample = str_c(t.exp, bio.rep, exp.plate,trt, sep = "_")) %>% 
  # blanks are not useful for this analysis
  filter(bio.rep!="blank") 

```

Look at data

```{r}   
   

p <- d.clean%>%
   pivot_longer(cols = c(veg.ml, spore.ml),
                names_to = "pop", values_to = "cell.ml" )%>%
  # mutate()
   ggplot(aes(x=bio.rep, y=cell.ml))+
   geom_point(aes(color=pop, shape=rep))+
   # facet_wrap(~interaction(time.hr,exp.plate))+
  facet_wrap(trt~time.hr+exp.plate)+
   theme_cowplot()+
   panel_border()+
   scale_y_log10()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
         strip.background = element_blank(),
         strip.text = element_text(size=8, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")))

ggsave(here("fig", "look-at-TRT-data.png"),p,
      width = 11, height = 8)
```
Mostly technical replicates are similar (=rep). Some that are off should mostly be solved by removing outliers.
Biological replication also look good.

# N samples
```{r}
d.clean%>%
  group_by(time.hr, exp.plate, trt) %>% 
  summarise(n=n()) %>% 
  # pivot_wider(names_from = exp.plate,values_from = n) %>% 
  ggplot(aes(time.hr, exp.plate)) +
  geom_tile(aes(fill = n), color = "white")+
  theme_bw()
```

dilutions
```{r}
d.clean%>%
  ggplot(aes(time.hr, volume.nL/1000))+
  # ggplot(aes(time.hr, total.events/1000))+
  geom_point(aes(shape = exp.plate,color=dilution))+
  # facet_wrap(~exp.plate, ncol = 2)+
  theme_bw()
```


# Replication outliers
I have made 3 technical replicate measurements in the flow-cytometer for each sample. There are some measurements which are obviously way off, e.g. in "WED12" samples some wells did not get SYBR to to pipetting error. To get rid of these I will choose from each triplicate the 2 points that are in best agreement and remove the 3rd point. There is a function to idenify such a point in pacakge 'outliers'
> outlier {outliers} Finds value with largest difference between it and sample mean, which can be an outlier.

> logical: if set to TRUE, gives vector of logical values, and possible outlier position is marked by TRUE

I will apply the outlier filtering on the number of spores (spore/mL)

```{r id outliers, echo=FALSE}


d2 <- tibble()
for (i in unique(d.clean$unq.sample)){
  
  tmp <- d.clean %>% 
    filter (unq.sample ==i)
  

  if (nrow(tmp)<3){
    tmp$out.lier <- FALSE
  } else if (sum(tmp$spore.ml)<300){
    tmp <- tmp %>% 
    mutate(out.lier = outlier(veg.ml, logical = T)) 
  } else {
    tmp <- tmp %>% 
    mutate(out.lier = outlier(spore.ml, logical = T)) 
  }

   d2 <- bind_rows(d2,tmp)
    
}



p <- d2%>%
   ggplot(aes(x=bio.rep, y=veg.ml))+
    geom_jitter(aes(color=out.lier), shape=21, width = 0.2, height = 0)+
  facet_wrap(trt~time.hr+exp.plate)+
   theme_cowplot()+
   panel_border()+
   scale_y_log10()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
         strip.background = element_blank(),
         strip.text = element_text(size=8, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")))

ggsave(here("fig", "TRT_Veg_ouliers.png"),p,
      width = 11, height = 8)

p <- d2%>%
   ggplot(aes(x=bio.rep, y=spore.ml))+
   geom_jitter(aes(color=out.lier), shape=21, width = 0.2, height = 0)+
  facet_wrap(trt~time.hr+exp.plate)+
   theme_cowplot()+
   panel_border()+
   scale_y_log10()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
         strip.background = element_blank(),
         strip.text = element_text(size=8, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")))

ggsave(here("fig", "TRT_Spore_ouliers.png"),p,
      width = 11, height = 8)

# remove outliers
d3 <- d2 %>% 
  filter(!out.lier)
```


Add T0 to all treatments (measured pre treatment)
```{r}
d3 <- d3 %>%
  select(bio.rep, trt) %>% 
  distinct() %>% 
  filter(trt != "preTrt") %>% 
  left_join(., d3 %>% filter(time.hr == 0) %>% select(-trt)) %>% 
  bind_rows(., d3 %>% filter(trt != "preTrt")) 
            

```
```{r}
d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml),names_to = "pop", values_to = "n.ml") %>% 
  
  group_by(time.hr, trt, bio.rep, pop) %>% 
  summarise(n.ml = mean(n.ml)) %>% 
  # mutate(trt = fct_relevel(trt, "LYSATE",after = 3)) %>% 

  ggplot(aes(time.hr, n.ml, color = pop)) +
  geom_vline(xintercept = 1.5, color = "red", linetype = 2)+
  geom_line()+
  geom_point(shape = 21, fill = "white")+
  facet_grid( trt %>% fct_relevel("LYSATE",after = 3) ~bio.rep )+
  facet_grid(bio.rep ~ trt %>% fct_relevel("LYSATE",after = 3))+
    theme_bw(base_size = 14)+
  panel_border(color = "black")+
  theme(strip.background = element_blank())+
  coord_cartesian(ylim = c(1e5,NA))+
  scale_y_log10()
  
```
### export data
```{r}
d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml) %>% round() ,
            spore.ml = mean(spore.ml)%>% round()) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 
  write_csv(here("data", "clean_counts.csv"))
```



```{r}
d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml),names_to = "pop", values_to = "n.ml") %>% 
  
  group_by(time.hr, trt, pop) %>% 
  summarise(m = mean(n.ml), v = sd(n.ml)) %>% 
  filter(pop == "spore.ml") %>% 

  ggplot(aes(time.hr, m, color = trt)) +
  geom_vline(xintercept = 1.5, color = "red", linetype = 2)+
  geom_line(size = 1)+
  geom_pointrange(aes(ymax = m+v, ymin = m-v),
                  shape = 21, fill = "white")+
    theme_bw()+
  panel_border(color = "black")+
  theme(strip.background = element_blank())+
  coord_cartesian(ylim = c(1e0,NA))+
  scale_y_log10()+
  ylab("Total cells/mL suspension")+
  xlab("Time after spotting on DSM agar (hr)")
  
```
```{r}
d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml),names_to = "pop", values_to = "n.ml") %>% 
  filter(pop == "spore.ml") %>% 

  ggplot(aes(time.hr, n.ml, color = trt)) +
  geom_vline(xintercept = 1.5, color = "red", linetype = 2)+
  geom_smooth(size = 1)+
  geom_point(shape = 21, fill = "white")+
    theme_bw()+
  panel_border(color = "black")+
  theme(strip.background = element_blank())+
  coord_cartesian(ylim = c(1e0,NA))+
  scale_y_log10()+
  ylab("Spres cells/mL suspension")+
  xlab("Time after spotting on DSM agar (hr)")+ 
  facet_wrap(~trt)
  
```
```{r}
d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml),names_to = "pop", values_to = "n.ml") %>% 
  filter(pop == "spore.ml") %>% 

  ggplot(aes(time.hr, n.ml, color = trt)) +
  geom_vline(xintercept = 1.5, color = "red", linetype = 2)+
  geom_smooth(size = 1)+
  # geom_point(shape = 21, fill = "white")+
    theme_bw()+
  panel_border(color = "black")+
  theme(strip.background = element_blank())+
  coord_cartesian(ylim = c(1e0,NA))+
  scale_y_log10()+
  ylab("Spores/mL suspension")+
  xlab("Time after spotting on DSM agar (hr)")
  # facet_wrap(~trt)
  
```
```{r}

d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt, bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml),names_to = "pop", values_to = "n.ml") %>% 
  
    group_by(time.hr, trt,pop) %>%
  summarise(m = mean(n.ml), v= sd(n.ml) ) %>% 

  
  ggplot(aes(time.hr, m))+
  geom_pointrange(aes(ymin = m-v, ymax = m+v))+
  stat_smooth(se = F)+
  # scale_y_log10()+

  theme_cowplot()+
  panel_border(color = "black")+
  facet_grid(trt~pop)+
  theme(strip.background = element_blank())
  

  
```
```{r}

d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt,bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml) %>% round(), spore.ml = mean(spore.ml)%>% round()) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml, 
          perc.spore =spore.ml*100/cell.ml ) %>% 
  
  # write_csv(here("data","Sporulation_by_Time_FCM.csv"))

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml, perc.spore),
               names_to = "pop", values_to = "n.ml") %>% 
  
  # mutate(n.ml = if_else(pop == "perc.spore", n.ml, log10(n.ml))) %>% 
  group_by(trt,pop) %>% 
    ggplot(aes(time.hr, n.ml))+
  geom_point(aes(color = bio.rep), shape = 21)+
  stat_smooth()+
  facet_grid(pop ~ trt, scales = "free_y")+
  theme_cowplot()+
  panel_border(color = "black")+
   theme(strip.background = element_blank())
  
```

```{r}

d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt,bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml, 
          perc.spore =spore.ml*100/cell.ml ) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml, perc.spore),
               names_to = "pop", values_to = "n.ml") %>% 
  group_by(pop) %>% 
  mutate(rel.abundance = n.ml/max(n.ml)) %>% 
    ggplot(aes(time.hr, rel.abundance))+
  geom_point(aes(color = exp.plate))+
  # stat_smooth()+
    stat_smooth(formula = "y ~ x", method = "glm", 
              method.args = list(family="quasibinomial"), se = T) +
  facet_grid(trt~pop, scales = "free_y")+

  theme_cowplot()+
  panel_border(color = "black")+
   theme(strip.background = element_blank())
  
```
```{r}
# Follwing https://www.ipb.pt/~vcadavez/websiteVC/tutorial/rcode/2019-04-28-gompertzmodel/

Gompertz <- function(x, y0, ymax, k, lag){
      result <- y0 + (ymax -y0)*exp(-exp(k*(lag-x)/(ymax-y0) + 1) )
      return(result)
      }

test.data <- 
  d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt,bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml, 
          perc.spore =spore.ml*100/cell.ml ) %>% 
  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml, perc.spore),
               names_to = "pop", values_to = "n.ml")   %>% 
  mutate(logN = ifelse(n.ml < 1, 1, log10(n.ml))) 

d.stats <- tibble()  
for (i_trt in unique(test.data$trt)){
  for(i_br in unique(test.data$bio.rep)){
    # for (i_pop in unique(test.data$pop)){
    for (i_pop in "spore.ml"){
      
      tmp <- 
        test.data %>% 
        filter(trt == i_trt) %>% 
        filter(pop == i_pop) %>%
        filter(bio.rep == i_br) %>%
        select(Time = time.hr,exp.plate,bio.rep, logN) %>% 
        as.data.frame()
    tryCatch({  
      Gomp1 <-  nls(logN ~ Gompertz(Time, y0, ymax, k, lag),
               data = tmp,
               start = list( ymax=7, k=0.5, lag=5),
              control = list(maxiter = 500))


      # extract model parameters
      coefs <- coef(Gomp1)
      
      d.stats <- tibble(
          trt = i_trt,
          pop = i_pop,
          bio.rep = i_br,
          # y0   = coefs[[1]],
          ymax = coefs[[1]],
          k   = coefs[[1]],
          lag = coefs[[3]]
      ) %>% 
        bind_rows(d.stats,.)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

    }
  }
}

filter(trt == "BUFFER") %>% 
  filter(pop == "spore.ml") %>% 
   filter(bio.rep == "br2") %>%
  mutate(logN = ifelse(n.ml < 1, 1, log10(n.ml))) %>% 
  select(Time = time.hr,exp.plate,bio.rep, logN) %>% 
  as.data.frame()
    # plot(test.data$Time, test.data$logN)






plot(1:20,Gompertz(1:20,y0 = y0, ymax = ymax, k = k, lag=lag))
```


```{r}
# Follwing https://www.ipb.pt/~vcadavez/websiteVC/tutorial/rcode/2019-04-28-gompertzmodel/

test.data <- 
  d3 %>% 
  group_by(sample.date,exp.plate,t.exp,time.hr, trt,bio.rep) %>% 
  summarise(veg.ml = mean(veg.ml), spore.ml = mean(spore.ml)) %>% 
  ungroup() %>% 
   mutate(cell.ml = veg.ml + spore.ml, 
          perc.spore =spore.ml*100/cell.ml ) %>% 

  pivot_longer(cols = c(cell.ml, spore.ml, veg.ml, perc.spore),
               names_to = "pop", values_to = "n.ml") %>%  
  filter(trt == "BUFFER") %>% 
  filter(pop == "spore.ml") %>% 
   filter(bio.rep == "br2") %>%
  mutate(logN = ifelse(n.ml < 1, 1, log10(n.ml))) %>% 
  select(Time = time.hr,exp.plate,bio.rep, logN) %>% 
  as.data.frame()
    # plot(test.data$Time, test.data$logN)

Gompertz <- function(x, y0, ymax, k, lag){
      result <- y0 + (ymax -y0)*exp(-exp(k*(lag-x)/(ymax-y0) + 1) )
      return(result)
      }


Gomp1 <-  nls(logN ~ Gompertz(Time, y0, ymax, k, lag),
               data = test.data,
               start = list(y0=0.15, ymax=7, k=0.5, lag=5),
              control = list(maxiter = 500))


# ectract model parameters
coefs <- coef(Gomp1)
coefs
 #       y0      ymax         k       lag 
 # 1.569712  8.792401 12.205638 10.578883 
y0   <- coefs[[1]]
ymax <- coefs[[2]]
k   <- coefs[[3]]
lag <- coefs[[4]]

plot(1:20,Gompertz(1:20,y0 = y0, ymax = ymax, k = k, lag=lag))
```

