---
title: "virospore annulus (Magalie) -- VIROSPORES only"
author: "Joy O'Brien"
date: "2025-04-24"
output: html_document
---
Joshua's feedback from the first figure--things to remember to do this time: 
- Do not draw means or standard deviations 
- Connect lines for each plaque
- Maybe add poisson counting statistics as SEM for each of the points

```{r setup, include=FALSE}
setwd("~/GitHub/VirosporeQuant/Magalie_virospore_quant")
```

```{r}
library(readxl)
library(dplyr)
library(tidyr)  
library(vegan)
library(ggplot2)
library(mgcv)

# Read in data
viro_quant <- read_excel("~/GitHub/VirosporeQuant/Magalie_virospore_quant/Data/Annulus_Quant/20250329_2253_viromultiquant_magalie.xlsx", sheet = "raw_data")

```

0. Set plot theme
```{r}
library(ggplot2)
theme_set(theme_classic() +
  theme(axis.title = element_text(size = 16),
        axis.title.x = element_text(margin = margin(t = 15, b = 15)),
        axis.title.y = element_text(margin = margin(l = 15, r = 15)),
        axis.text = element_text(size = 13),
        axis.text.x = element_text(margin = margin(t = 5)),
        axis.text.y = element_text(margin = margin(r = 5)),
        #axis.line.x = element_line(linewidth = 1),
        #axis.line.y = element_line(linewidth = 1),
        axis.ticks.y.right = element_line(linewidth = 1),
        axis.text.y.right = element_text(size = 13, margin = margin(l = 5)),
        axis.line.y.right = element_line(linewidth = 0.1),
        axis.ticks.x.top = element_line(linewidth = 1),
        axis.text.x.top = element_text(size = 13, margin = margin(b = 5)),
        axis.line.x.top = element_line(linewidth = 0.1),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_line(linewidth = 1),
        axis.ticks.y = element_line(linewidth = 1),
        axis.ticks.length = unit(.1, "in"),
        panel.border = element_rect(color = "grey20", fill = NA, linewidth = 1.5),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 14),
        strip.background = element_blank()
        )) 
```

1. Calculate VPFU/mL
```{r}
# We want to calculate VPFU, CFU and PFU of the original suspension which is # of PFU or CFU / vol plated x total dilution factor

# VPFU only (for now)
viro_quant$VPFU_mL_1 <- viro_quant$VPFU_10e1 / (0.1 * (10^-1)) # using data from the 10^-1 dilution, disregarding the 10^-2 data
#viro_quant$VPFU_mL_2 <- viro_quant$VPFU_10e2 / (0.1 * (10^-2))
```

1B. Calculate CFU/mL and PFU/mL per stock solution for each sample
```{r}
viro_quant$VPFU_stock <- viro_quant$VPFU_mL_1 * 1.5 # 1.5 mL is the volume that the sample was deposited into
```
1C. Calulate the amount of virospores per cm^2 area
```{r}
# For virospores
viro_quant$Virospores_cm2 <- viro_quant$VPFU_stock / (0.00785) # this is the area of the pipette tip in cm ^2
```

2. Plotting counts
```{r}
viro_quant.new <- viro_quant[, c("Plaque", "Sample", "Region", "Virospores_cm2", "Distance_from_PFU_center_cm")]

viro_quant.long <- viro_quant.new %>%
  pivot_longer(cols = c(Virospores_cm2), 
               names_to = "variable", 
               values_to = "value") 
```

Plot with shading
```{r}
# Define shading regions
region_shading <- data.frame(
  Region = c("Center", "Annulus", "Lawn"),
  xmin = c(0, 0.05, 0.1),   # Starting x-value for each region
  xmax = c(0.05, 0.1, 0.15)  # Ending x-value for each region
)
region_shading$Region <- factor(region_shading$Region, levels = c("Center", "Annulus", "Lawn"))

# Plot with facet wrap
ggplot(viro_quant.long, aes(x = Distance_from_PFU_center_cm, y = value, color = Plaque, group = Plaque)) +
  # Add shaded regions
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            alpha = 0.3) +
  # Line and point layers
  geom_point(size = 2) + 
  geom_line() + 
  # Custom fill colors for shaded regions
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  coord_fixed(ratio = 0.1) + # Setting distance between x axis ticks
  # Labels 
  labs(title = "Virospore Plot", 
       x = "Distance from PFU center (cm)", 
       y = "Virospores per cm^2") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```

3. Averaging across plaques and plotting
```{r}
library(dplyr)

# First, assign regions based on distance if not already present
library(dplyr)

# Summarize across plaques
viro_quant_avg <- viro_quant.long %>%
  group_by(Distance_from_PFU_center_cm) %>%  # Use Region if needed instead of distance
  summarise(
    mean_virospores = mean(value, na.rm = TRUE),
    n = n(),  # number of plaques (should be 3)
    SEM = sd(value, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
head(viro_quant_avg)


library(ggplot2)

ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region),
            alpha = 0.3) +
  geom_point(color = "darkgreen", size = 3) +
  geom_line(color = "darkgreen") +
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "darkgreen") +
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +
  scale_y_log10() +
   #coord_fixed(ratio = 0.1) + # Setting distance between x axis ticks
  labs(
    title = "Average Virospore Counts Across 3 Plaques",
    x = "Distance from PFU center (cm)",
    y = "Mean Virospores per cm²"
  ) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


```
4. Plotting with Joshua's previous feedback in mind and using the color in Magalie data
```{r}
library(ggplot2)

ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region),
            alpha = 0.3) +
  geom_point(color = "#94f293", size = 3) +
  # Adding smooth lines
  geom_smooth(method = "loess", color = "#94f293", se = FALSE, size = 1) + 
  # geom_line(color = "darkgreen")  # You can remove this if you want only smooth lines
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "black") +
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +
  scale_y_log10() +
  labs(
    title = "",
    x = "Distance from PFU center (cm)",
    y = "Mean Virospores per cm²"
  ) +
  coord_fixed(ratio = 0.1) + # Setting distance between x axis ticks
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```
4A. Plotting the above again but this time with no shading in the background for the regions of the plaque, instead dotted lines will signify regions 
```{r}
library(ggplot2)

ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_point(color = "#94f293", size = 3) +
  # Adding smooth lines
  geom_smooth(method = "loess", color = "#94f293", se = FALSE, size = 1) + 
  # geom_line(color = "darkgreen")  # You can remove this if you want only smooth lines
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "black") +
  # Adding in vertical lines
  geom_vline(xintercept = c(0.05, 0.1),
             linetype = "dashed", color = "darkgray", size = 0.8) +
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +
  scale_y_log10() +
  labs(
    title = "",
    x = "Distance from PFU center (cm)",
    y = "Mean Virospores per cm²"
  ) +
  theme_minimal() +
  coord_fixed(ratio = 0.1)  # Setting distance between x axis ticks
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
4B. Attempting to make it look more like the other figures in the manuscript

```{r}
library(ggplot2)

ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_point(color = "#94f293", size = 3) +
  geom_smooth(method = "loess", color = "#94f293", se = FALSE, size = 1) +
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "black") +
  geom_vline(xintercept = c(0.05, 0.1),
             linetype = "dashed", color = "darkgray", size = 0.8) +
  scale_y_log10() +
  labs(
    title = "",
    x = "Distance from PFU center (cm)",
    y = "Virospores per cm²"
  ) +
  coord_fixed(ratio = 0.1)



```


```{r}
library(ggplot2)

ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_point(color = "#94f293", size = 3) +
  geom_smooth(method = "loess", color = "#94f293", se = FALSE, size = 1) +
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "black") +
  geom_vline(xintercept = c(0.05, 0.1),
             linetype = "dashed", color = "darkgray", size = 0.8) +
  #scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +
  scale_y_log10() +
  labs(
    title = "",
    x = "Distance from PFU center (cm)",
    y = "Virospores per cm²"
  ) +  #theme(axis.ticks.y.right = element_line(linewidth = 1)) +
  #scale_y_continuous(sec.axis = sec_axis(~., labels = NULL)) +
  coord_fixed(ratio = 0.1)

    #panel.border = element_rect(color = "black", fill = NA, size = 1),
    #axis.line = element_line(color = "black"), 
    #plot.title = element_text(hjust = 0.5, vjust = 0.5)
  #)
```
Editing code for the following: 
x-axis: “distance from center (cm)” to match figures in Magalie
y-axis: typically don’t say “mean”. This will be denoted in caption
y-axis: “Virospore density (#/cm^2)”
plot titles: not included in publications
To match Magalie, we might consider plotting standard deviations and means as above on left.

```{r}
library(ggplot2)

ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_point(color = "#94f293", size = 3) +
  geom_smooth(method = "loess", color = "#94f293", se = FALSE, size = 1) + 
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "black") +
  geom_vline(xintercept = c(0.05, 0.1),
             linetype = "dashed", color = "darkgray", size = 0.8) +
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +
  scale_y_log10() +
  labs(
    title = "", # no need to include plot titles in publications
    x = "Distance from center (cm)",
    y = "Virospores (#/cm²)"
  ) +
  theme_classic() +    # <<-- use classic theme instead of minimal
  coord_fixed(ratio = 0.1) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    axis.line = element_line(color = "black"), 
    plot.title = element_text(hjust = 0.5, vjust = 0.5)
  )

```



Trying to get the tick marks to appear
```{r}
viro.quant <- ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = mean_virospores)) +
  geom_point(color = "#94f293", size = 3) +
  geom_smooth(method = "loess", color = "#94f293", se = FALSE, size = 1) +
  geom_errorbar(aes(ymin = mean_virospores - SEM, ymax = mean_virospores + SEM),
                width = 0.005, color = "black") +
  geom_vline(xintercept = c(0.05, 0.1),
             linetype = "dashed", color = "darkgray", size = 0.8) +
  scale_x_continuous(
    name = "Distance from PFU center (cm)",
    sec.axis = dup_axis(name = NULL, labels = NULL)  # Adds top axis with same ticks
  ) +
  scale_y_log10(
    name = "Virospores (#/cm²)",
    sec.axis = dup_axis(name = NULL, labels = NULL)  # Adds right axis with same ticks
  ) +
  coord_fixed(ratio = 0.1) +
  labs(title = "")
ggsave("~/GitHub/danger.signal/Magalie_virospore_quant/Figures/virospore_plot.png", viro.quant, width = 6, height = 4)
```



