---
title: "Virospore annulus multi-quant for Magalie"
author: "Joy O'Brien"
date: "2025-03-22"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/GitHub/VirosporeQuant/Magalie_virospore_quant")
```

```{r}
# Load Library
library(readxl)
library(dplyr)
library(tidyr)  
library(vegan)
library(ggplot2)
library(mgcv)

# Read in data
viro_quant <- read_excel("~/GitHub/VirosporeQuant/Magalie_virospore_quant/Data/Annulus_Quant/20250329_2253_viromultiquant_magalie.xlsx", sheet = "raw_data")
```

# Calculate CFU, PFUs, and VPFUs per mL
```{r}
# Calculate CFU and PFU of the original suspension which is # of PFU or CFU / vol plated x total dilution factor
#viro_quant$Free_PFU_mL_1 <- viro_quant$Free_PFU_10e1 / (0.1 * (10^-1))
viro_quant$Free_PFU_mL_2 <- viro_quant$Free_PFU_10e2/ (0.1 * (10^-2))

# Uninfected spores
#viro_quant$Spores_CFU_mL_1 <- viro_quant$Spores_CFU_10e1 / (0.1 * (10^-1))
viro_quant$Spores_CFU_mL_2 <- viro_quant$Spores_CFU_10e2 / (0.1 * (10^-2))

# VPFU
viro_quant$VPFU_mL_1 <- viro_quant$VPFU_10e1 / (0.1 * (10^-1))
viro_quant$VPFU_mL_2 <- viro_quant$VPFU_10e2 / (0.1 * (10^-2))


```

1B. Calculate CFU/mL and PFU/mL per stock solution for each sample
```{r}
viro_quant$Free_PFU_stock <- viro_quant$Free_PFU_mL_2 * 1.5
#viro_quant$Free_PFU_5stock <- viro_quant$Free_PFU_mL_5rep * 1.5
#viro_quant$Free_PFU_7stock <- viro_quant$Free_PFU_mL_7rep * 1.5

viro_quant$Spores_CFU_stock <- viro_quant$Spores_CFU_mL_2 * 1.5
#viro_quant$Spores_CFU_5stock <- viro_quant$Spores_CFU_mL_5rep * 1.5
#viro_quant$Spores_CFU_7stock <- viro_quant$Spores_CFU_mL_7rep * 1.5

viro_quant$VPFU_stock <- viro_quant$VPFU_mL_1 * 1.5
```

Extra: Calculate entrapment efficiency according to Gabiatti et al 2018 --> although this may not be a good idea for plate assays (?)
```{r}
# Trying to calculate entrapment efficiency via Gabiatti 
viro_quant$e.e <- viro_quant$VPFU_stock / (viro_quant$VPFU_stock + viro_quant$Spores_CFU_stock) * 100

# Averaging the entrapment effieicny among the replicate plaques so that I can report an average
avg_efficiency <- viro_quant %>%
  group_by(Region) %>%
  summarize(MeanEfficiency = mean(e.e, na.rm = TRUE),
            SD = sd(e.e, na.rm = TRUE),
            N = n())

library(ggplot2)

ggplot(avg_efficiency, aes(x = Sample, y = MeanEfficiency)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Entrapment Efficiency",
       x = "Sample",
       y = "Entrapment Efficiency (Entrapped / Total PFU)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Okay getting back on track...

1C. Calculate the amount of Free PFU and Spores CFU per cm^2 area
```{r}
# Calculate per area for free phage
viro_quant$Free_PFU_cm2 <- viro_quant$Free_PFU_stock / (0.00785)
#viro_quant$Free_PFU_5_cm2 <- viro_quant$Free_PFU_5stock / (0.00785)
#viro_quant$Free_PFU_7_cm2 <- viro_quant$Free_PFU_7stock / (0.00785)

# For free spores
viro_quant$Spores_CFU_cm2 <- viro_quant$Spores_CFU_stock / (0.00785)
#viro_quant$Spores_CFU_5_cm2 <- viro_quant$Spores_CFU_5stock / (0.00785)
#viro_quant$Spores_CFU_7_cm2 <- viro_quant$Spores_CFU_7stock / (0.00785)

# For virospores
viro_quant$Virospores_cm2 <- viro_quant$VPFU_stock / (0.00785)
```

Subset data to plot 
```{r}
# Plotting counts
viro_quant.new <- viro_quant[, c("Plaque", "Sample", "Region", "Virospores_cm2", "Distance_from_PFU_center_cm", "Free_PFU_cm2", "Spores_CFU_cm2")]

# Calculate total spores with VPFU and the averages (if needed)
viro_quant.new$Spores_total <- viro_quant.new$Spores_CFU_cm2 + viro_quant.new$Virospores_cm2
```


Preparing to plot data
```{r}
# Convert data to long format
viro_quant.long <- viro_quant.new %>%
  pivot_longer(cols = c(Virospores_cm2, Free_PFU_cm2, Spores_CFU_cm2), 
               names_to = "variable", 
               values_to = "value") 

# Define shading regions of where the annulus, kill zone, and lawn is
region_shading <- data.frame(
  Region = c("Center", "Annulus", "Lawn"),
  xmin = c(0, 0.05, 0.1),   # Starting x-value for each region
  xmax = c(0.05, 0.1, 0.15)  # Ending x-value for each region
)
region_shading$Region <- factor(region_shading$Region, levels = c("Center", "Annulus", "Lawn"))

# Plot with facet wrap
ggplot(viro_quant.long, aes(x = Distance_from_PFU_center_cm, y = value, color = variable, group = variable)) +
  # Add shaded regions
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            alpha = 0.3) +
  # Line and point layers
  geom_line() +  
  geom_point(size = 2) + 
  # Custom fill colors for shaded regions
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  # Labels 
  labs(title = "Virospore Multi-Quant Plot", 
       x = "Distance from PFU center (cm)", 
       y = "Virospores, Free Phage, Uninfected Spores per cm^2", 
       color = "Measurement") +
  theme_classic() +
  facet_wrap(~Plaque)  # Facet by Plaque

```
Plotting with different colors and altering the x axis
```{r}

ggplot(viro_quant.long, aes(x = Distance_from_PFU_center_cm, y = value, color = variable, group = variable)) +
  # Add shaded regions
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            alpha = 0.3) +
  # Line and point layers
  geom_line(linewidth = 1.2) +  
  geom_point(size = 2) + 
  # Custom colors for lines with correct label placement
  scale_color_manual(
    values = c("Virospores_cm2" = "#457B9D", "Free_PFU_cm2" = "#E63946", "Spores_CFU_cm2" = "#2A9D8F"),
    labels = c("Virospores_cm2" = "Virospores", "Free_PFU_cm2" = "Free phage", "Spores_CFU_cm2" = "Uninfected spores")  # Correct placement of labels
  ) +  
  # Custom fill colors for shaded regions
  scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  # Labels 
  labs(title = "", 
       x = "Distance from PFU center (cm)", 
       y = "Virospores, Free Phage, Uninfected Spores per cm²", 
       color = "Measurement") +
  theme_classic(base_size = 12) +
  
  # Ensure x-axis values overlap properly
  theme(axis.text.x = element_text(angle = 50, vjust = 0.5)) +
  # Expand the plotting area to prevent cut-off points
  coord_cartesian(clip = "off") +
  # Facet by Plaque (duplicate removed)
  facet_wrap(~Plaque)

```


2. Plot all samples as individuals
```{r}
viro_quant.long.all <- viro_quant.new %>%
  pivot_longer(cols = c(Virospores_cm2),
   names_to = "variable", 
               values_to = "value")
# Plot
ggplot(viro_quant.long.all, aes(x = Distance_from_PFU_center_cm, y = value, color = variable, group = variable)) +
  # Add shaded regions using geom_rect()
  geom_rect(data = region_shading, inherit.aes = FALSE, 
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            alpha = 0.3) +
  # Line and point layers
  geom_line(linewidth = 1.2) +  
  geom_point(aes(color = variable), size = 2) + 
  # Custom fill colors for shaded regions
  scale_fill_manual(values = c("Center" = "lightblue", "Annulus" = "lightgreen", "Lawn" = "lightpink")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  # Labels 
  labs(title = "Virospore Multi-Quant Plot Plaque A", 
       x = "Distance from PFU center (cm)", 
       y = "Virospores, Free Phage, Uninfected Spores per cm^2", 
       color = "Measurement") +
  theme_minimal()
```

Plotting the averge data of the 3 replicate PFUs
```{r}
library(dplyr)
library(ggplot2)

# Compute the average across plaques
viro_quant_avg <- viro_quant.long %>%
  group_by(Distance_from_PFU_center_cm, variable) %>%
  summarize(value = mean(value), .groups = "drop")

# Define shading regions
#region_shading <- data.frame(
  #Region = c("Center", "Annulus", "Lawn"),
  #xmin = c(0, 0.05, 0.1),   # Starting x-value for each region
  #xmax = c(0.05, 0.1, 0.15)  # Ending x-value for each region
#)
#region_shading$Region <- factor(region_shading$Region, levels = c("Center", "Annulus", "Lawn"))

# Plot averaged data
ggplot(viro_quant_avg, aes(x = Distance_from_PFU_center_cm, y = value, color = variable, group = variable)) +
  # Add shaded regions
  #geom_rect(data = region_shading, inherit.aes = FALSE, 
            #aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = Region), 
            #alpha = 0.3) +
  # Line and point layers + 
  geom_line() +
  geom_point(size = 2) + 
  geom_vline(xintercept = c(0.05, 0.1), 
             linetype = "dashed", color = "darkgray", size = 0.8) +
  # Custom fill colors for shaded regions
  #scale_fill_manual(values = c("Center" = "white", "Annulus" = "lightgray", "Lawn" = "darkgray")) +  
  # Log transformation for y-axis
  scale_y_log10() +
  coord_fixed(ratio = 0.1) +
  # Labels 
  labs(title = "", 
       x = "Distance from center (cm)", 
       y = "Virospores, Free Phage, Uninfected Spores (#/cm²)", 
       color = "Measurement") +
  theme_classic()

```



