---
title: "Spo0A:GFP Plate reader analysis for Magalie"
author: "Joy O'Brien"
date: "2025-05-07"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/GitHub/danger.signal/3.plate_reader")
```

1. Load data
```{r}
gfp <- read.csv("~/GitHub/danger.signal/3.plate_reader/GrowthCurves/data/20250213_spo0agfp_gfp_reads.csv")
treats.gfp <- read.csv("~/GitHub/danger.signal/3.plate_reader/GrowthCurves/data/20250213_treatments_spo0A_gfp.csv")

```

1A. Prep/clean data for plotting RFU over time
```{r}
# Remove data from empty wells (20250213 plate)
gfp <- subset(gfp, select = -c(A2, A3, A5, A6, A8, A9, A11, A12, B1, B2, B3, B4, B5, B6, B7, B8, B9, B10, B11, B12, C1, C2, C3, C4, C5, C6, C7, C8, C9, C10, C11, C12, D2, D3, D5, D6, D8, D9, D11, D12, E1, E2, E3, E4, E5, E6, E7, E8, E9, E10, E11, E12, F1, F2, F3, F4, F5, F6, F7, F8, F9, F10, F11, F12, G2, G3, G5, G6, G8, G9, G11, G12, H1, H2, H3, H4, H5, H6, H7, H8, H9, H10, H11, H12))


# For the 20240213 data where two different media are used, we need to do it twice
DSM_columns_to_adjust <- c("A1","A7","A10","D1","D4","D7")
cDSM_columns_to_adjust <- c("A4", "D10", "G1", "G4", "G7", "G10")

gfp <- gfp
gfp[DSM_columns_to_adjust] <- sweep(gfp[DSM_columns_to_adjust], 1, gfp$A1, "-")

gfp <- gfp
gfp[cDSM_columns_to_adjust] <- sweep(gfp[cDSM_columns_to_adjust], 1, gfp$A4, "-")



# Print the adjusted data frame
print(gfp)

# Change times to numeric
gfp$Time <- as.numeric(sub("^(\\d+):(\\d+).*", "\\1.\\2", gfp $Time))
head(gfp, header = T)




```
1B. Rescale the data
```{r}
# Rescale the data before transforming
rescale_minmax <- function(x, lower = 1, upper = 10) {
  return(lower + (x - min(x)) / (max(x) - min(x)) * (upper - lower))
}
# Trying Z-score normalization/rescaling
#zscore_normalization <- function(x) {
 # return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
#}

# Specify columns to scale
columns_to_scale <- c("A7", "A10", "D1", "D4", "D7", "D10", "G1", "G4", "G7", "G10")
gfp[columns_to_scale] <- lapply(gfp[columns_to_scale], rescale_minmax)
```



1C. Rename columns based on sample type
```{r}
# Rename the columns based on the sample instead of well number
library(dplyr)
library(tidyr)
gfp <- gfp %>%
  rename(DSM_blank = A1, cDSM_blank = A4, GFP_DSM_control_R1 = A7, GFP_DSM_control_R2 = A10, GFP_DSM_BBphage_R1= D1, GFP_DSM_BBphage_R2= D4, GFP_DSM_BBphage_R3 = D7, GFP_cDSM_control_R1 = D10, GFP_cDSM_control_R2
= G1, GFP_cDSM_Bbphage_R1 = G4, GFP_cDSM_Bbphage_R2 = G7, GFP_cDSM_Bbphage_R3= G10)
```

1D. Convert to long format
```{r}
# Pivot data to longer format
gfp_long <- gfp %>%
  pivot_longer(
    cols = -c(Time, Temp.),  # keep Time and Temp. as-is
    names_to = "Sample",     # name for the new column holding old column names
    values_to = "GFP"        # name for the new column holding values
  )
# View
head(gfp_long)
```




```{r}


```

2. Plotting cDSM and DSM together (first look)
```{r}
library(ggplot2)
# Subset the data 
ggplot(gfp_long, aes(x = Time, y = GFP, color = Sample)) +
  #geom_line() +
  geom_smooth(span = 0.4, alpha = 0.2) +
  #geom_smooth(span, se = TRUE, alpha = 0.2) +
  #geom_point() +
  #scale_y_log10() +
  labs(
    title = "",
    x = "Time (hours)",
    y = "GFP (RFU)",
    color = "Sample"
  ) +
    theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    panel.grid.major = element_blank(),  # removes major grid lines
    panel.grid.minor = element_blank(),  # removes minor grid lines
    panel.background = element_blank(),  # removes the background color
    plot.background = element_blank(),  # removes background around the plot
    theme_minimal()  # minimal theme
  )
```
3. Subsetting data based on media type (cDSM, DSM)
```{r}
# Remove any cDSM result (subsetting cDSM data and DSM)
gfp_long_DSM <- gfp_long %>%
  filter(!grepl("cDSM", Sample))

# View the result
head(gfp_long_DSM)

```

3A. Plotting just the DSM samples 
```{r}
library(ggplot2)
# Subset the data 
ggplot(gfp_long_DSM, aes(x = Time, y = GFP, color = Sample)) +
  #geom_line() +
  geom_smooth(span = 0.4, alpha = 0.2) +
  #geom_smooth(span, se = TRUE, alpha = 0.2) +
  #geom_point() +
  #scale_y_log10() +
  labs(
    title = "",
    x = "Time (hours)",
    y = "GFP (RFU)",
    color = "Sample"
  ) +
    theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    panel.grid.major = element_blank(),  # removes major grid lines
    panel.grid.minor = element_blank(),  # removes minor grid lines
    panel.background = element_blank(),  # removes the background color
    plot.background = element_blank(),  # removes background around the plot
    theme_minimal()  # minimal theme
  ) + 
   annotate("text", x = 20, y = 10, label = "Rm-ANOVA p = 0.6", color = "black", size = 5)
```
Editing figure for publication
```{r}


```

```{r}


```







3B. Add a treatment column and remove blanks from the DSM df
```{r}
# Data needs to be in long format and contain columns that specify treatment, time, and RFU
library(stringr)

gfp_DSM <- gfp_long_DSM %>%
  mutate(treatment = case_when(
    str_detect(Sample, regex("phage", ignore_case = TRUE)) ~ "Phage particles",
    str_detect(Sample, regex("control", ignore_case = TRUE)) ~ "Control",
    str_detect(Sample, regex("blank", ignore_case = TRUE)) ~ "Blank",
    TRUE ~ "other"  # fallback category if none match
  ))
# Remove blank samples
#gfp_DSM <- gfp_DSM %>%
  #filter(treatment != "blank")
```

Editing figure for publication/to look like the rest

```{r}

ggplot(gfp_DSM, aes(x = Time, y = GFP, color = treatment)) +
  geom_smooth(span = 0.4, alpha = 0.2) +
  labs(
    title = "",
    x = "Time (hours)",
    y = "GFP (RFU)",
    color = "Treatment"
  ) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  ) + 
  annotate("text", x = 20, y = 10, label = "Rm-ANOVA p = 0.6", color = "black", size = 5)

```







3C. Run repeated-measures ANOVA on DSM samples
```{r}
# Sample is a factor identifying the repeated measures (between-subject factor)
# Time is the repeated measure (within subject factor)
# GFP is the response variable
str(gfp_DSM)
gfp_DSM$Sample <- as.factor(gfp_DSM$Sample)

aov_result_DSM <- aov(GFP ~ treatment * Time + Error(Sample / Time), data = gfp_DSM)
summary(aov_result_DSM)

# AOV results no significant difference in the effect of treatment on samples p value = 0.603
```

Extra:
```{r}

# In the event that things were significant (as we found when blanks were included)
#Looking at post-hoc since we know that blanks cause false significance 
library(emmeans)

emmeans_result_DSM <- emmeans(aov_result_DSM, ~ treatment | Time)
pairs(emmeans_result_DSM)
```


3D. Plotting cDSM samples
```{r}

# Subset samples that include cDSM
gfp_cDSM <- gfp_long[grepl("cDSM", gfp_long$Sample), ]


ggplot(gfp_cDSM, aes(x = Time, y = GFP, color = Sample)) +
  #geom_line() +
  geom_smooth(span = 0.4, alpha = 0.2) +
  #geom_smooth(span, se = TRUE, alpha = 0.2) +
  #geom_point() +
  #scale_y_log10() +
  labs(
    title = "",
    x = "Time (hours)",
    y = "GFP (RFU)",
    color = "Sample"
  ) +
    theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    panel.grid.major = element_blank(),  # removes major grid lines
    panel.grid.minor = element_blank(),  # removes minor grid lines
    panel.background = element_blank(),  # removes the background color
    plot.background = element_blank(),  # removes background around the plot
    theme_minimal()  # minimal theme
  ) +
  annotate("text", x = 20, y = 10, label = "Rm-ANOVA p = 0.417", color = "black", size = 5)



```
3E. Add treatment column and remove blanks from the cDSM
```{r}
# Data needs to be in long format and contain columns that specify treatment, time, and RFU
library(stringr)

gfp_cDSM <- gfp_cDSM %>%
  mutate(treatment = case_when(
    str_detect(Sample, regex("phage", ignore_case = TRUE)) ~ "phage",
    str_detect(Sample, regex("control", ignore_case = TRUE)) ~ "control",
    str_detect(Sample, regex("blank", ignore_case = TRUE)) ~ "blank",
    TRUE ~ "other"  # fallback category if none match
  ))

# Remove blanks from the dataset because they are causing false significant results
# Remove blank samples
gfp_cDSM <- gfp_cDSM %>%
  filter(treatment != "blank")
```

3F. Run repeated measures anova on the cDSM samples 
```{r}
# Sample is a factor identifying the repeated measures (between-subject factor)
# Time is the repeated measure (within subject factor)
# GFP is the response variable
str(gfp_cDSM)
gfp_cDSM$Sample <- as.factor(gfp_cDSM$Sample)

aov_result_cDSM <- aov(GFP ~ treatment * Time + Error(Sample / Time), data = gfp_cDSM)
summary(aov_result_cDSM)
```
Extra:
```{r}
# If things were significant and we did a post hoc follow up 
  
library(emmeans)

emmeans_result_cDSM <- emmeans(aov_result_cDSM, ~ treatment | Time)
pairs(emmeans_result_cDSM)
```

4. Moving on to compare Gompertz estimated onset parameters (cDSM first) -- come back and check why the lag times are weird
```{r}
# Since this data was previously run through Gompertz already, we just need to call the params
# Loading cDSM params data
library(readr)
gfp.cDSM.controls.1parms <- read.csv("~/GitHub/danger.signal/3.plate_reader/GrowthCurves/output/gfp.cDSM.controls.1.parms.txt",header = TRUE)
gfp.cDSM.bbphage.1parms <- read.csv("~/GitHub/danger.signal/3.plate_reader/GrowthCurves/output/gfp.cDSM.BBphage.1.parms.txt", header = TRUE)

#cDSM params aren't making sense to me, need to go back and check what happened 
```




5. Moving on to compare Gompertz estimated onset parameters (DSM)
```{r}
gfp.DSM.controls.1parms <- read.csv("~/GitHub/danger.signal/3.plate_reader/GrowthCurves/output/gfp.DSM.controls.1.parms.txt",header = TRUE)
gfp.DSM.bbphage.1parms <- read.csv("~/GitHub/danger.signal/3.plate_reader/GrowthCurves/output/gfp.DSM.BBphage.1.parms.txt", header = TRUE)
```

5A. We want to plot and show the following: Mean onset time, error bars, raw data points (can be done with ggplot jitter I think)
```{r}
# First combine the treatment and control data together
# Assume df_exp and df_ctrl are your experimental and control datasets
gfp.DSM.bbphage.1parms$treatment <- "Phage"
gfp.DSM.controls.1parms$treatment <- "Control"

# Combine both
gfp_DSM_lcombo <- rbind(gfp.DSM.bbphage.1parms, gfp.DSM.controls.1parms)

# Plot
ggplot(gfp_DSM_lcombo, aes(x = treatment, y = L)) +
  geom_boxplot(outlier.shape = NA, fill = "#94f293", alpha = 0.5) +
  #stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "red") +
  #stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "red") +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  labs(
    x = "Treatment",
    y = "GFP Onset time (hours)",
    title = "DSM Spo0A:GFP onset time comparison"
  ) +
  theme_mininal()
  
  
  
# Plot for publication 
ggplot(gfp_DSM_lcombo, aes(x = treatment, y = L)) +
  geom_boxplot(outlier.shape = NA, fill = "#94f293", alpha = 0.5) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  labs(
    x = "Treatment",
    y = "GFP Onset time (hours)",
    title = "DSM Spo0A:GFP onset time comparison"
  ) +
  theme_minimal() +  # Add minimal theme first
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  ) +
  annotate("text", x = 2, y = 23, label = "ANOVA p = 0.549", color = "black", size = 5)

```
Running ANOVA on this data
```{r}
anova.DSM <- aov(L ~ treatment, data = gfp_DSM_lcombo)
summary(anova.DSM)

```



```

