---
title: "D.S. flow cytometry analysis for Magalie"
author: "Joy O'Brien"
date: "2025-05-06"
output: html_document
---
Goal: Create figures that show null results for danger signals for Magalie et al from Daniel Schwartz's flow cytometry data (obtained from flow-spore-time repo)
```{r setup, include=FALSE}
setwd("~/GitHub/danger.signal/1.flow_cytometry")
```
0. Set theme
```{r}
library(paletteer)
library(ggplot2)
theme_set(theme_classic() +
  theme(axis.title = element_text(size = 16),
        axis.title.x = element_text(margin = margin(t = 15, b = 15)),
        axis.title.y = element_text(margin = margin(l = 15, r = 15)),
        axis.text = element_text(size = 13),
        axis.text.x = element_text(margin = margin(t = 5)),
        axis.text.y = element_text(margin = margin(r = 5)),
        #axis.line.x = element_line(linewidth = 1),
        #axis.line.y = element_line(linewidth = 1),
        axis.ticks.y.right = element_line(linewidth = 1),
        axis.text.y.right = element_text(size = 13, margin = margin(l = 5)),
        axis.line.y.right = element_line(linewidth = 0.1),
        axis.ticks.x.top = element_line(linewidth = 1),
        axis.text.x.top = element_text(size = 13, margin = margin(b = 5)),
        axis.line.x.top = element_line(linewidth = 0.1),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_line(linewidth = 1),
        axis.ticks.y = element_line(linewidth = 1),
        axis.ticks.length = unit(.1, "in"),
        panel.border = element_rect(color = "grey20", fill = NA, linewidth = 1.5),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 14),
        strip.background = element_blank()
        )) 
colors <-paletteer_d("ggthemes::Miller_Stone")
```

1. Load data
```{r}
flow.spore <- read.csv("~/GitHub/danger.signal/1.flow_cytometry/Data/clean_counts.csv")
```

2. Set up data to create figure of % spores and time
```{r}
library(ggplot2)
# Subset the data 
ggplot(flow.spore, aes(x = time.hr, y = spore.ml, color = trt)) +
  #geom_line() +
  geom_smooth(span = 0.4, alpha = 0.2) +
  #geom_smooth(span, se = TRUE, alpha = 0.2) +
  #geom_point() +
  labs(
    title = "",
    x = "Time (hours)",
    y = "Spores per mL",
    color = "Treatment"
  ) #+
  #theme(
  #legend.position = "right",  # adjust legend position
    #theme_minimal()
  #)





```
3.Setting up data to show means + SEM
```{r}
library(dplyr)
fc.summary <- flow.spore %>%
  group_by(trt, time.hr) %>%
  summarise(
    mean_spores = mean(spore.ml, na.rm = TRUE),
    sem_spores = sd(spore.ml, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

library(dplyr)

fc.summary <- fc.summary %>%
  mutate(trt = recode(trt,
                          "DSM" = "Fresh media",
                          "LYSATE" = "Phage lysate",
                          "SPENT" = "Spent media", 
                          "BUFFER" = "Buffer"))
```


3A. Plotting for publication 
```{r}

#Setting colors 
colors <- c(
  "Spent media" = "#F47942FF",
  "Phage lysate" = "#638B66FF",
  "Fresh media" = "#849DB1FF",
  "Buffer" = "#FBB04EFF"
)

fc.time.plot <- ggplot(fc.summary, aes(x = time.hr, y = mean_spores, color = trt)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = mean_spores - sem_spores, ymax = mean_spores + sem_spores),
                width = 0.2, color = "black") +
  labs(
    title = "",
    x = "Time (h)",
    y = "Spores (#/mL)",
    color = "Treatment"
  ) +
  scale_x_continuous(
    limits = c(10,40),
    breaks = c(10, 15, 20, 25, 30),
    sec.axis = dup_axis(name = NULL, labels = NULL)
  ) +
  scale_color_manual(values = colors) +
  scale_y_continuous(sec.axis = dup_axis(name = NULL, labels = NULL)) + 
  #theme()
  #axis.line.y.right = element_line(linewidth = 0.1), 
  #axis.line.x.top = element_line(linewidth = 0.1)
  # theme(legend.position = "none") +
  theme(legend.position = "none") +
  annotate("text", x = 36, y = 800000000, label = "Spent media", color =  "#F47942FF", size = 3) +
  annotate("text", x = 36, y = 880000000, label = "Phage lysate", color =  "#638B66FF", size = 3) +
  annotate("text", x = 36, y = 1255000000, label = "Fresh media", color = "#849DB1FF", size = 3) +
  annotate("text", x = 36, y = 950000000, label = "Buffer", color = "#FBB04EFF", size = 3)

print(fc.time.plot)  

```

4. Repeated measures ANOVA
```{r}

aov_result <- aov(spore.ml ~ trt * time.hr + Error(bio.rep/time.hr), data = flow.spore)
summary(aov_result)

```
Given the outputs of the Rm-ANOVA we know: 
- spore count changes significantly over time regardless of treatment (expected)
- there is no statistically significant in spore yield due to treatment (null result)
- there is a statistically significant treatment and time interaction




4A.Figure that includes RM-ANOVA p-value
```{r}
ggplot(flow.spore, aes(x = time.hr, y = spore.ml, color = trt)) +
  #geom_point() +
  geom_smooth(span = 0.4, alpha = 0.2) + 
  labs(
    title = "",
    x = "Time (hours)",
    y = "Spores per mL",
    color = "Treatment"
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    panel.grid.major = element_blank(),  # removes major grid lines
    panel.grid.minor = element_blank(),  # removes minor grid lines
    panel.background = element_blank(),  # removes the background color
    plot.background = element_blank(),  # removes background around the plot
    theme_minimal()  # minimal theme
  ) +
  annotate("text", x = 20, y = 1500000000, label = "Rm-ANOVA p = 0.064", color = "black", size = 5)
```

5. Set up data to estimate onset of sporulation (lag time) via Gompertz
```{r setup}
rm(list=ls())
setwd("~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves")
knitr::opts_knit$set(root.dir = "~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves")
```

```{r}
# Load packages and functions
require("png")
require("dplyr")
require("grid")
require("gtools")
require("nlme")
require("MuMIn")
require("bbmle")
source("~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves/bin/modified_Gomp_diagnostic3.R")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) 100*( sd(x)/mean(x))
```

5A. Load data
```{r}
flow.spore <- read.csv("~/GitHub/danger.signal/1.flow_cytometry/Data/clean_counts.csv")

# Change time to numeric 

```


5B. Wrangle data before GrowthCurver analysis

```{r}
library(dplyr)
library(tidyr)
# Group by treatment

fc.buffer <- flow.spore %>%
  filter(trt %in% c("BUFFER"))

fc.DSM <- flow.spore %>%
  filter(trt %in% ("DSM"))

fc.lysate <- flow.spore %>%
  filter(trt %in% ("LYSATE"))

fc.spent <- flow.spore %>%
  filter(trt %in% ("SPENT") )

# Convert to wide
fc.buffer.wide<- fc.buffer %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.DSM.wide <- fc.DSM %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.lysate.wide <- fc.lysate %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.spent.wide <- flow.spore %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

# Change time 
fc.buffer.wide <- fc.buffer.wide %>%
  rename(Time = time.hr)

fc.DSM.wide <- fc.DSM.wide %>%
  rename(Time = time.hr)

fc.lysate.wide <- fc.lysate.wide %>%
  rename(Time = time.hr)

fc.spent.wide <- fc.spent.wide %>%
  rename(Time = time.hr)


# Collapse multiple rows per time point via dplyr
fc.buffer.wide <- fc.buffer.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

fc.DSM.wide <- fc.DSM.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

fc.lysate.wide <- fc.lysate.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

fc.spent.wide <- fc.spent.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

# Remove data that is not needed
fc.buffer.wide<- fc.buffer.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

fc.DSM.wide<- fc.DSM.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

fc.lysate.wide<- fc.lysate.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

fc.spent.wide<- fc.spent.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)


```

5C. Setting up to run Gompertz via GrowthCurver
```{r}
fc.buffer.wide.df <- as.data.frame(fc.buffer.wide)
fc.DSM.wide.df <- as.data.frame(fc.DSM.wide)
fc.lysate.wide.df <- as.data.frame(fc.lysate.wide)
fc.spent.wide.df <- as.data.frame(fc.spent.wide)

library(growthcurver)
# Time should not be in 00:00 format

buffer.plate <- fc.buffer.wide.df
  names(buffer.plate)

DSM.plate <- fc.DSM.wide.df
  names(DSM.plate)

lysate.plate <- fc.lysate.wide.df
  names(lysate.plate)

spent.plate <- fc.spent.wide.df
  names(spent.plate)

# Run the growth curve analysis
buffer.plate.summary <-SummarizeGrowthByPlate(plate = buffer.plate)
head(buffer.plate.summary)

DSM.plate.summary <-SummarizeGrowthByPlate(plate = DSM.plate)
head(DSM.plate.summary)

lysate.plate.summary <-SummarizeGrowthByPlate(plate = lysate.plate)
head(lysate.plate.summary)

spent.plate.summary <- SummarizeGrowthByPlate(plate = spent.plate)
head(spent.plate.summary)

# Calculate lag time for all
buffer.plate.summary$lag_time <- buffer.plate.summary$t_mid - (1 / buffer.plate.summary$r)
DSM.plate.summary$lag_time <- DSM.plate.summary$t_mid - (1 / DSM.plate.summary$r)
lysate.plate.summary$lag_time <- lysate.plate.summary$t_mid - (1 / lysate.plate.summary$r)
spent.plate.summary$lag_time <- spent.plate.summary$t_mid - (1 / spent.plate.summary$r)


#Combine into one df but first add treatments
buffer.plate.summary$Treatment <- "Buffer"
DSM.plate.summary$Treatment <- "DSM"
spent.plate.summary$Treatment <- "Spent"
lysate.plate.summary$Treatment <- "Lysate"


lag_data <- bind_rows(
  buffer.plate.summary,
  DSM.plate.summary,
  spent.plate.summary,
  lysate.plate.summary
)
```
5D.Initial plot
```{r}
# Plot for publication 
ggplot(lag_data, aes(x = Treatment, y = lag_time)) +
  geom_boxplot(outlier.shape = NA, fill = "#94f293", alpha = 0.5) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  labs(
    x = "Treatment",
    y = "Sporulation onset (h)",
    title = ""
  ) +
  #theme_minimal() +  # Add minimal theme first
  #theme(
    # panel.border = element_rect(color = "black", fill = NA, size = 1),
    # panel.grid.major = element_blank(),
    # panel.grid.minor = element_blank(),
    # panel.background = element_blank(),
    # plot.background = element_blank()
  #) #+
  annotate("text", x = 3.7, y = 17, label = "ANOVA p-value = 0.731", color = "black", size = 5)
```
5E. Anova of sporulation onset times and treatment 
```{r}
# ANOVA
anova.lag <- aov(lag_time ~ Treatment, data = lag_data)
summary(anova.lag)
```

5F. Plotting for publication
```{r}
library(legendry)
# Rename the treatments for publication 
lag_data <- lag_data %>%
  mutate(Treatment = recode(Treatment,
                          "DSM" = "Fresh media",
                          "Lysate" = "Phage lysate",
                          "Spent" = "Spent media"))

lag_data$Treatment <- factor(lag_data$Treatment, levels =c("Buffer", "Fresh media", "Spent media", "Phage lysate"))
library(legendry)

# Okay the above code adds the top ticks but does not preserve the bottom x axis 
fc.box.plot <- ggplot(lag_data, aes(x = Treatment, y = lag_time)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  scale_x_discrete(labels = c(
    "Buffer",
    "Fresh\nmedia",
    "Spent\nmedia",
    "Phage\nlysate"
  )) +
  labs(
    x = "Treatment",
    y = "Sporulation onset (h)",
    title = ""
  ) +
  scale_y_continuous(sec.axis = sec_axis(~., labels = NULL)) +
  annotate("text", x = 3, y = 18, label = "ANOVA p-value = 0.7", color = "black") +
  #theme(
    #axis.ticks.x.top = element_line(),       # Show top ticks
    #axis.line.x.top = element_line(),        # Draw top axis line
    #axis.text.x.top = element_blank(),       # Optional: remove top labels
    #axis.title.x.top = element_blank()       # Optional: remove top title
  #) +
  guides(x.sec = legendry::guide_axis_base(key = key_map(data = lag_data, aesthetic = unique(Treatment), label = c("", "", "", ""))))

print(fc.box.plot)
```
6. Plot the time series plot and box plot together
```{r}
library(ggplot2)
library(cowplot)
#library(magick)
#plate.img <- image_read("~/GitHub/danger.signal/2.square_plate/Data/23_h-1.tif")

#Convert image to cowplot object
# Step 2: Convert image to a cowplot object
#img_plot <- ggdraw() + draw_image(plate.img)
empty.plot <- ggdraw()
# Okay actually we will add the image later, create an empty plot 
# Step 3: Combine all plots with the image on the left
fc.combined_plots <- ggdraw() +
  draw_plot(empty.plot, x = 0,   y = 0, width = 1/3, height = 1) +  # Left image
  draw_plot(fc.time.plot,  x = 1/3, y = 0, width = 1/3, height = 1) +  # Center plot
  draw_plot(fc.box.plot,   x = 2/3, y = 0, width = 1/3, height = 1) +  # Right plot
  theme(plot.background = element_rect(fill= "white", color = NA)) +
  draw_plot_label(label = c("(A)", "(B)", "(C)"), size = 16,
                  x = c(0.05, 0.38, 0.71), y = c(1, 1, 1))

print(fc.combined_plots)
ggsave("~/GitHub/danger.signal/1.flow_cytometry/Figures/fc.combined_plotswspace.png", fc.combined_plots, width = 12, height = 4)
```

Extra: Trying to run modified Gompertz out of GrowthCurves3B
```{r}
# Separate data out by experimental plate 
# fc.buffer.exp1 <- fc.buffer %>%
#   filter(exp.plate == "plt1")
# 
# fc.buffer.exp2 <- fc.buffer %>%
#   filter(exp.plate == "plt2")

# Rename the sample names so we know the treatment


# Transform to wide format so we can run the model
library(tidyr)

fc.buffer.wide<- fc.buffer %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.buffer.wide <- fc.buffer.wide %>%
  rename(buffer_br1 = br1)

fc.buffer.wide <- fc.buffer.wide.exp1 %>%
 rename(buffer_br2 = br2)

fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br3 = br3)

# Rename some columns 
fc.buffer.wide <- fc.buffer.wide %>%
  rename(Time = time.hr)
# Separate out by experimental plate too?
fc.buffer.exp1 <- fc.buffer %>%
  filter(exp.plate == "plt1")

fc.buffer.exp2 <- fc.buffer %>%
  filter(exp.plate == "plt2")

# Average across biological replicates?

# Rename the sample names so we know the treatment
fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br1 = br1)

fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br2 = br2)

fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br3 = br3)

# Remove unneeded data
fc.buffer.wide.exp.1.min <- fc.buffer.wide.exp1 %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

# Collapse multiple rows per time point via dplyr
fc.buffer.wide.clean.exp1 <- fc.buffer.wide.exp.1.min %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

# Remove samples without any variation (Check to see if this is causing the error below)
fc.buffer.wide.clean.exp1 <- fc.buffer.wide.clean.exp1 %>%
  select(Time, where(~ sd(., na.rm = TRUE) > 0))

```


Extra: Attempt to run modified Gompertz
```{r}
setwd("~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves/bin")
fc.buffer.1 <- growth.modGomp(input = fc.buffer.wide.clean.exp1, output.name = "fc.buffer.exp1.1.parms",
                 synergy = F, temp = F, smooth = T, trim = T)
```
# Error from running Gompertz 
Error in ts(x) : 'ts' object must have one or more observations


```{r}
fc.DSM.1 <- growth.modGomp(input = fc.DSM,output.name = "fc.DSM.1.parms",
                  synergy = F, temp = F, smooth = T, trim = T)
fc.lysate.1  <- growth.modGomp(input = fc.lysate.1, output.name = "fc.lysate.1.parms",
                 synergy = F, temp = F, smooth = T, trim = T)
fc.spent.1 <- growth.modGomp(input = fc.spent.1, output.name = "fc.spent.1.parms",
               synergy = F, temp = F, smooth = T, trim = T)

```





