---
title: "D.S. flow cytometry analysis for Magalie"
author: "Joy O'Brien"
date: "2025-05-06"
output: html_document
---
Goal: Create figures that show null results for danger signals for Magalie et al from Daniel Schwartz's flow cytometry data (obtained from flow-spore-time repo)
```{r setup, include=FALSE}
setwd("~/GitHub/danger.signal/1.flow_cytometry")
```

1. Load data
```{r}
flow.spore <- read.csv("~/GitHub/danger.signal/1.flow_cytometry/Data/clean_counts.csv")
```

2. Set up data to create figure of % spores and time
```{r}
library(ggplot2)
# Subset the data 
ggplot(flow.spore, aes(x = time.hr, y = spore.ml, color = trt)) +
  #geom_line() +
  geom_smooth(span = 0.4, alpha = 0.2) +
  #geom_smooth(span, se = TRUE, alpha = 0.2) +
  #geom_point() +
  labs(
    title = "",
    x = "Time (hours)",
    y = "Spores per mL",
    color = "Treatment"
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    theme_minimal()
  )





```


2A. Repeated measures ANOVA
```{r}

aov_result <- aov(spore.ml ~ trt * time.hr + Error(bio.rep/time.hr), data = flow.spore)
summary(aov_result)

```
Given the outputs of the Rm-ANOVA we know: 
- spore count changes significantly over time regardless of treatment (expected)
- there is no statistically significant in spore yield due to treatment (null result)
- there is a statistically significant treatment and time interaction




Figure that includes RM-ANOVA p-value
```{r}
ggplot(flow.spore, aes(x = time.hr, y = spore.ml, color = trt)) +
  #geom_point() +
  geom_smooth(span = 0.4, alpha = 0.2) + 
  labs(
    title = "",
    x = "Time (hours)",
    y = "Spores per mL",
    color = "Treatment"
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.position = "right",  # adjust legend position
    panel.grid.major = element_blank(),  # removes major grid lines
    panel.grid.minor = element_blank(),  # removes minor grid lines
    panel.background = element_blank(),  # removes the background color
    plot.background = element_blank(),  # removes background around the plot
    theme_minimal()  # minimal theme
  ) +
  annotate("text", x = 20, y = 1500000000, label = "Rm-ANOVA p = 0.064", color = "black", size = 5)
```

3. Set up data to estimate onset of sporulation (lag time) via Gompertz
```{r setup}
rm(list=ls())
setwd("~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves")
knitr::opts_knit$set(root.dir = "~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves")
```


```{r}
# Load packages and functions
require("png")
require("dplyr")
require("grid")
require("gtools")
require("nlme")
require("MuMIn")
require("bbmle")
source("~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves/bin/modified_Gomp_diagnostic3.R")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) 100*( sd(x)/mean(x))
```

3A. Load data
```{r}
flow.spore <- read.csv("~/GitHub/danger.signal/1.flow_cytometry/Data/clean_counts.csv")

# Change time to numeric 

```


3B. Wrangle data before GrowthCurver analysis

```{r}
# Group by treatment

fc.buffer <- flow.spore %>%
  filter(trt %in% c("BUFFER"))

fc.DSM <- flow.spore %>%
  filter(trt %in% ("DSM"))

fc.lysate <- flow.spore %>%
  filter(trt %in% ("LYSATE"))

fc.spent <- flow.spore %>%
  filter(trt %in% ("SPENT") )

# Convert to wide
fc.buffer.wide<- fc.buffer %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.DSM.wide <- fc.DSM %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.lysate.wide <- fc.lysate %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.spent.wide <- flow.spore %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

# Change time 
fc.buffer.wide <- fc.buffer.wide %>%
  rename(Time = time.hr)

fc.DSM.wide <- fc.DSM.wide %>%
  rename(Time = time.hr)

fc.lysate.wide <- fc.lysate.wide %>%
  rename(Time = time.hr)

fc.spent.wide <- fc.spent.wide %>%
  rename(Time = time.hr)


# Collapse multiple rows per time point via dplyr
fc.buffer.wide <- fc.buffer.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

fc.DSM.wide <- fc.DSM.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

fc.lysate.wide <- fc.lysate.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

fc.spent.wide <- fc.spent.wide %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

# Remove data that is not needed
fc.buffer.wide<- fc.buffer.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

fc.DSM.wide<- fc.DSM.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

fc.lysate.wide<- fc.lysate.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

fc.spent.wide<- fc.spent.wide %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)


```

Setting up to run Gompertz via GrowthCurver
```{r}
fc.buffer.wide.df <- as.data.frame(fc.buffer.wide)
fc.DSM.wide.df <- as.data.frame(fc.DSM.wide)
fc.lysate.wide.df <- as.data.frame(fc.lysate.wide)
fc.spent.wide.df <- as.data.frame(fc.spent.wide)

library(growthcurver)
# Time should not be in 00:00 format

buffer.plate <- fc.buffer.wide.df
  names(buffer.plate)

DSM.plate <- fc.DSM.wide.df
  names(DSM.plate)

lysate.plate <- fc.lysate.wide.df
  names(lysate.plate)

spent.plate <- fc.spent.wide.df
  names(spent.plate)

# Run the growth curve analysis
buffer.plate.summary <-SummarizeGrowthByPlate(plate = buffer.plate)
head(buffer.plate.summary)

DSM.plate.summary <-SummarizeGrowthByPlate(plate = DSM.plate)
head(DSM.plate.summary)

lysate.plate.summary <-SummarizeGrowthByPlate(plate = lysate.plate)
head(lysate.plate.summary)

spent.plate.summary <- SummarizeGrowthByPlate(plate = spent.plate)
head(spent.plate.summary)

# Calculate lag time for all
buffer.plate.summary$lag_time <- buffer.plate.summary$t_mid - (1 / buffer.plate.summary$r)
DSM.plate.summary$lag_time <- DSM.plate.summary$t_mid - (1 / DSM.plate.summary$r)
lysate.plate.summary$lag_time <- lysate.plate.summary$t_mid - (1 / lysate.plate.summary$r)
spent.plate.summary$lag_time <- spent.plate.summary$t_mid - (1 / spent.plate.summary$r)



#Combine into one df but first add treatments
buffer.plate.summary$Treatment <- "Buffer"
DSM.plate.summary$Treatment <- "DSM"
spent.plate.summary$Treatment <- "Spent"
lysate.plate.summary$Treatment <- "Lysate"


lag_data <- bind_rows(
  buffer.plate.summary,
  DSM.plate.summary,
  spent.plate.summary,
  lysate.plate.summary
)
```
```{r}
# Plot for publication 
ggplot(lag_data, aes(x = Treatment, y = lag_time)) +
  geom_boxplot(outlier.shape = NA, fill = "#94f293", alpha = 0.5) +
  geom_jitter(width = 0.1, size = 2, alpha = 0.7) +
  labs(
    x = "Treatment",
    y = "GFP Onset time (hours)",
    title = ""
  ) +
  theme_minimal() +  # Add minimal theme first
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  ) +
  annotate("text", x = 2, y = 17, label = "ANOVA p = 0.731", color = "black", size = 5)
```
```{r}
# ANOVA
anova.lag <- aov(lag_time ~ Treatment, data = lag_data)
summary(anova.lag)
```


```{r}

```






3B. Wrangle data (trying this on one subset of data before doing all treatments)
```{r}
# Separate data out by experimental plate 
# fc.buffer.exp1 <- fc.buffer %>%
#   filter(exp.plate == "plt1")
# 
# fc.buffer.exp2 <- fc.buffer %>%
#   filter(exp.plate == "plt2")

# Rename the sample names so we know the treatment


# Transform to wide format so we can run the model
library(tidyr)

fc.buffer.wide<- fc.buffer %>%
  pivot_wider(names_from = bio.rep, values_from = spore.ml)

fc.buffer.wide <- fc.buffer.wide %>%
  rename(buffer_br1 = br1)

fc.buffer.wide <- fc.buffer.wide.exp1 %>%
 rename(buffer_br2 = br2)

fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br3 = br3)

# Rename some columns 
fc.buffer.wide <- fc.buffer.wide %>%
  rename(Time = time.hr)
# Separate out by experimental plate too?
fc.buffer.exp1 <- fc.buffer %>%
  filter(exp.plate == "plt1")

fc.buffer.exp2 <- fc.buffer %>%
  filter(exp.plate == "plt2")

# Average across biological replicates?

# Rename the sample names so we know the treatment
fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br1 = br1)

fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br2 = br2)

fc.buffer.wide.exp1 <- fc.buffer.wide.exp1 %>%
  rename(buffer_br3 = br3)

# Remove unneeded data
fc.buffer.wide.exp.1.min <- fc.buffer.wide.exp1 %>%
  select(-sample.date, -exp.plate, -t.exp, -veg.ml, -cell.ml, -trt)

# Collapse multiple rows per time point via dplyr
fc.buffer.wide.clean.exp1 <- fc.buffer.wide.exp.1.min %>%
  group_by(Time) %>%
  summarise(across(everything(), ~ first(na.omit(.))), .groups = "drop")

# Remove samples without any variation (Check to see if this is causing the error below)
fc.buffer.wide.clean.exp1 <- fc.buffer.wide.clean.exp1 %>%
  select(Time, where(~ sd(., na.rm = TRUE) > 0))

```


3D. Run Gompertz
```{r}
setwd("~/GitHub/danger.signal/1.flow_cytometry/GrowthCurves/bin")
fc.buffer.1 <- growth.modGomp(input = fc.buffer.wide.clean.exp1, output.name = "fc.buffer.exp1.1.parms",
                 synergy = F, temp = F, smooth = T, trim = T)
```
# Error from running Gompertz 
Error in ts(x) : 'ts' object must have one or more observations


```{r}
fc.DSM.1 <- growth.modGomp(input = fc.DSM,output.name = "fc.DSM.1.parms",
                  synergy = F, temp = F, smooth = T, trim = T)
fc.lysate.1  <- growth.modGomp(input = fc.lysate.1, output.name = "fc.lysate.1.parms",
                 synergy = F, temp = F, smooth = T, trim = T)
fc.spent.1 <- growth.modGomp(input = fc.spent.1, output.name = "fc.spent.1.parms",
               synergy = F, temp = F, smooth = T, trim = T)

```





